\subsection{Energy--Frequency Relation}
  \label{subsec:energy-frequency-solitons}

  Within effective descriptions of Cosmochrony, the energy associated with a
  particle-like excitation is linked to a characteristic internal spectral scale of
  the corresponding projected configuration.
  This scale quantifies how strongly the configuration resists admissible relaxation
  ordering: configurations associated with higher characteristic frequencies correspond
  to more tightly constrained projected structures and therefore encode a greater
  effective resistance to relaxation.

  This correspondence provides an effective interpretation of the relation
  \begin{equation}
    E \propto \nu ,
  \end{equation}
  in which energy measures the degree of effective relaxation resistance associated
  with a projected configuration, while the frequency $\nu$ characterizes the spectral
  scale of its internal organization.
  The frequency should not be interpreted as a literal oscillation with respect to a
  fundamental time parameter.
  It acquires an operational temporal meaning only within effective geometric regimes,
  where a notion of time emerges through projected ordering.

  Within this framework, Planck's constant appears as an effective proportionality
  factor relating energy and frequency.
  Its apparent universality reflects the robustness of the spectral scales governing
  admissible projected configurations in the current relaxation epoch, rather than the
  postulation of a fundamental quantization constant at the level of the $\chi$
  substrate.

  In this sense, the energy--frequency relation expresses a kinematic correspondence
  between effective structural resistance and spectral organization within projected
  descriptions.
  A more explicit realization of this correspondence, in the context of radiation and
  photon-like projected excitations, is presented in
  Sec.~\ref{subsec:energy-frequency-radiation}.
