\subsection{Fermions and Bosons}
  \label{subsec:fermions-and-bosons}

  Within the Cosmochrony framework, particle statistics do not arise from fundamental
  quantization rules, canonical commutation relations, or postulated spin variables.
  They emerge at the level of effective descriptions from the topological structure
  of admissible projected configurations.

  Distinct classes of particle-like projected configurations are characterized by how
  their internal configuration space responds to continuous rotations within the space
  of admissible descriptions.
  Certain configurations require a $4\pi$ rotation to return to an equivalent
  projected description, while others are already $2\pi$-periodic.
  The former give rise to fermion-like behavior, whereas the latter correspond to
  boson-like excitations.

  This distinction reflects a topological obstruction in the space of admissible
  projected configurations rather than a symmetry principle imposed at the fundamental
  level.
  In effective geometric regimes, $4\pi$-periodic configurations may be described as
  possessing twisted or non-orientable internal structure, while $2\pi$-periodic
  configurations correspond to orientable ones.
  These characterizations are descriptive and do not imply the existence of a
  fundamental spatial manifold, intrinsic spin variables, or underlying rotational
  degrees of freedom.

  Within this perspective, the spin--statistics connection admits a natural qualitative
  interpretation.
  Fermionic and bosonic behavior reflects the topological classification of admissible
  projected configurations under continuous deformations, rather than the imposition
  of additional quantum postulates at the level of the $\chi$ substrate.

  As throughout this work, references to phase rotations, periodicity, or internal
  configuration space should be understood strictly within the effective descriptive
  framework.
  They characterize topological invariants of projected descriptions and their
  admissible transformations, not intrinsic attributes of the pre-geometric $\chi$
  substrate.
