\subsection{Antiparticles and CPT as an Admissibility Consistency Condition}
  \label{subsec:antiparticles-cpt}

  The emergence of antiparticles in decay processes is often associated, in standard
  quantum field theory, with CPT symmetry.
  In Cosmochrony, CPT is not imposed at a fundamental level.
  It is recovered as a consistency condition on admissible projected descriptions
  arising from the non-injective projection of the \(\chi\)-substrate.

  Certain projected configurations carry \emph{orientation-sensitive} structural
  invariants, such as chirality, phase winding, or relational orientation.
  These invariants are not defined at the fundamental \(\chi\) level; they acquire
  operational meaning only through projection and the effective geometric description.
  Under admissible factorization, such signed invariants cannot disappear.
  They must be redistributed among the resulting localized configurations.

  When a metastable projected configuration fragments, admissibility may therefore
  require the appearance of paired localized excitations carrying opposite orientations
  of the same invariant.
  At the effective level, these paired configurations are interpreted as
  particle--antiparticle pairs.
  Antiparticles are thus not entities propagating backward in a fundamental time, nor
  independent microscopic degrees of freedom postulated \emph{ab initio}.
  They are structural counterbalances required to preserve orientation-sensitive
  invariants during admissible reconfiguration.

  From this perspective, CPT symmetry reflects the invariance of admissibility under a
  combined reversal of (i) the relevant signed structural invariants (charge-conjugation
  at the effective level), (ii) effective spatial orientation (parity), and (iii) the
  effective ordering parameter associated with relational relaxation (time reversal in
  the emergent description).
  This combined transformation preserves the admissibility class of projected
  descriptions even when the individual transformations do not.

  In this sense, CPT is recovered as a robustness property of the projection structure:
  admissible projected descriptions remain consistent under the combined reversal of
  effective orientation data.
  A technical formulation of this invariance and its relation to signed structural
  invariants is provided in Appendix~\ref{subsec:structural-interpretation-of-cpt-symmetry}.
