\section{Glossary of Core Quantities and Notation}
  \label{appendix:glossary}

  This appendix summarizes the meaning and status of the main quantities used
  throughout the Cosmochrony framework.
  It is intended strictly as a reference guide and does not introduce
  new assumptions, dynamics, or physical postulates.

  \subsection{Fundamental and Effective Quantities}

    \paragraph{$\chi$ (Chi field).}
      The fundamental scalar quantity of the Cosmochrony framework.
      $\chi$ is not defined on a pre-existing spacetime manifold but constitutes a
      pre-geometric substrate whose monotonic relaxation provides an intrinsic
      ordering of physical processes.
      Localized, topologically stable configurations of $\chi$ correspond to
      particle-like excitations.

    \paragraph{$V(\chi)$ (Effective potential).}
      An effective, coarse-grained description used to model localization and
      stability properties of $\chi$ configurations.
      $V(\chi)$ is not assumed to be fundamental; its form may emerge from
      underlying discrete relaxation dynamics and is secondary to the spectral
      description of mass.

    \paragraph{$K_{ij}$ (Relaxation coupling).}
      Edge-dependent coupling coefficients defined on the relaxation network
      $G(V,E)$.
      $K_{ij}$ quantify the local resistance to relative variations of $\chi$
      between neighboring nodes and encode geometric and topological information
      of the network.
      They may depend on the local configuration of $\chi$ in effective descriptions.

\subsection{Derived Operators and Dimensionless Parameters}

  \paragraph{$G(V,E)$ (Relaxation network).}
    A discrete graph representing the underlying relational structure on which
    the $\chi$ field is defined at the fundamental level.
    Vertices correspond to elementary degrees of freedom, and edges encode
    relaxation couplings.

  \paragraph{$\Delta_G$ (Graph Laplacian / relaxation operator).}
    The discrete Laplace--Beltrami operator associated with the network $G(V,E)$
    and the couplings $K_{ij}$.
    It governs the stability and mode structure of localized $\chi$
    configurations.
    Its spectral properties play a central role in the quantitative
    characterization of inertial mass.

  \paragraph{$S$ (Gradient saturation parameter).}
    A dimensionless quantity defined as
    \begin{equation}
      S \equiv \frac{1}{c^2}\sum_{j\sim i} K_{ij}(\chi_i-\chi_j)^2 ,
    \end{equation}
    measuring the local density of $\chi$ gradients.
    The condition $S \leq 1$ ensures causal consistency and bounds the local
    relaxation rate of $\chi$.

  \paragraph{$\lambda_n$ (Spectral eigenvalues).}
    The eigenvalues of the linearized relaxation or stability operator acting on
    small fluctuations around a localized configuration.
    In effective wave descriptions, $\sqrt{\lambda_n}$ determines the inertial
    mass scale of particle-like excitations.

  \paragraph{$\Omega_\chi$ (Relaxation budget parameter).}
    A dimensionless global quantity characterizing the fraction of the total
    $\chi$ relaxation budget stored in spatial gradients.
    In cosmological regimes, $\Omega_\chi$ plays a role analogous to the matter
    density parameter in standard cosmology.

\subsection{Key Concepts}

\paragraph{Energy.}
  Energy is a conserved quantity associated with time-translation symmetry and
  the capacity to induce change.
  In Cosmochrony, energy is interpreted as a measure of resistance of
  $\chi$-field configurations to dynamical evolution.
  Standard conservation laws and empirical relations remain unaffected.

\paragraph{Decoherence.}
  In quantum mechanics, decoherence denotes the suppression of interference
  effects due to interaction with an environment.
  In Cosmochrony, decoherence is interpreted as the irreversible local
  deformation of the $\chi$ field induced by interaction, which destroys the
  phase correlations required for coherent superposition without altering the
  underlying structural configuration.

\paragraph{Fluctuations.}
  Fluctuations refer to stochastic variations of the $\chi$ field around a
  given configuration.
  They modulate the localization and timing of individual events without
  altering the underlying structural constraints imposed by the $\chi$
  topology.

\paragraph{Matter.}
  Matter conventionally refers to localized physical entities carrying mass and
  energy.
  Within Cosmochrony, matter corresponds to stable topological configurations
  of the $\chi$ field, whose persistence gives rise to particle-like behavior
  and inertial properties.

\paragraph{Measurement.}
  In standard quantum mechanics, a measurement refers to an interaction
  resulting in a definite outcome drawn from a probability distribution
  described by the wavefunction.
  In Cosmochrony, measurement is interpreted as a localized interaction that
  selects a specific manifestation of an underlying $\chi$-field fluctuation,
  without altering the global probabilistic structure associated with the
  system.
  This interpretation does not require a fundamental wavefunction collapse.

\paragraph{Probability.}
  Probabilities are not taken as primitive.
  They reflect stable structural constraints imposed by the local topology of
  the $\chi$ field, defining invariant patterns of allowed manifestations.
  Stochastic fluctuations of $\chi$ modulate this pattern at the event level.

\paragraph{Relaxation (of the $\chi$ field).}
  Relaxation refers to the intrinsic dynamical tendency of the $\chi$ field to
  continuously extend and reorganize its configuration under internal coupling
  constraints.
  This process is geometric and pre-thermodynamic in nature and does not
  correspond to dissipation or entropy maximization.

\paragraph{Schrödinger Equation.}
  An effective linear equation governing the evolution of quantum probability
  amplitudes.
  In Cosmochrony, the Schrödinger equation emerges as an approximate
  description of coherent, weak fluctuations of the $\chi$ field around a
  stable configuration.

\paragraph{Space--Time.}
  Spacetime is an emergent relational structure arising from large-scale
  configurations of the $\chi$ field.
  Its effective metric description remains valid at accessible scales.

\paragraph{Time.}
  Time is interpreted as an emergent parameter associated with the local rate
  of evolution of the $\chi$ field.
  Operational time measurements and relativistic predictions remain
  unchanged.

\paragraph{Uncertainty Principle.}
  The uncertainty principle arises from the fact that any interaction locally
  modifies the configuration of the $\chi$ field.
  Probing one observable necessarily affects complementary dynamical aspects
  of the field.

\paragraph{Wavefunction.}
  The wavefunction $\psi$ is an effective statistical representation emerging
  from the dynamics and topology of the underlying $\chi$ field.
  It is not a fundamental physical entity.

\paragraph{Wave--Particle Duality.}
  Wave--particle duality reflects interaction-induced changes in the local
  configuration of the $\chi$ field.
  The system remains fundamentally wave-like, while localized particle-like
  manifestations arise dynamically during interaction.
