\documentclass[11pt,a4paper]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{definition}{Definition}

\title{Spectral Admissibility and Non-Abelian Mode Resilience
in the $Q_8$ Cayley Graph under Bounded Flux}
\author{Working note -- Cosmochrony programme}
\date{\today}

\begin{document}
  \maketitle

  \begin{abstract}
    We establish the first analytically tractable instance of the spectral admissibility
    programme outlined in the Cosmochrony framework.
    On the undirected Cayley graph of the quaternion group $Q_8$, with generators
    $S = \{\pm i, \pm j, \pm k\}$, the local bounded-flux constraint
    $|\partial_t \chi_v| \leq c_\chi$ induces a hierarchy of maximal admissible
    effective amplitudes across irreducible spectral sectors.

    In the linearised dispersion regime, the non-abelian irreducible sector
    admits a maximal effective amplitude larger than the nontrivial abelian sectors
    by a factor of $\sqrt{4/3} \approx 1.155$.
    Resolving explicitly the basis dependence in the degenerate $\rho_4$ eigenspace,
    we show that this advantage holds for the physically relevant effective amplitude
    $A_n = a_n \|\psi_n\|_\infty$, but not for the raw amplitude $a_n$.

    We then perform the weakly non-linear modal reduction of the discrete
    Dirac--Born--Infeld (DBI) Lagrangian and derive the first amplitude-dependent
    correction to the dispersion relation, governed by the quartic shape factor
    $\kappa_n = \sum_v \psi_n(v)^4$.
    When expressed in terms of the effective amplitude, this correction depends
    only on the Laplacian eigenvalue $\lambda_n$, leaving the spectral hierarchy
    unchanged.

    Finally, solving the one-mode DBI dynamics exactly on the $Q_8$ Cayley graph,
    we show that the admissibility envelope
    \[
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}
    \]
    coincides with the exact saturation boundary of the DBI dynamics.
    The spectral hierarchy therefore remains structurally stable beyond the
    perturbative regime.

    This provides a minimal explicit demonstration that bounded substrate capacity
    can favour non-abelian relational modes purely through spectral admissibility,
    without invoking additional dynamical selection principles.

    More broadly, it suggests that non-abelian symmetry sectors may emerge
    as the dynamically most resilient components of the relational spectrum
    under bounded-flux constraints.
  \end{abstract}

  \section{Setting and Motivation}

    The Cosmochrony framework identifies the physical spectrum with the residue of two
    successive constraints acting on the relational operator $\mathcal{L}_\chi$:
    \begin{enumerate}
      \item \textbf{Topological discretisation}: the fibre structure of the non-injective
      projection $\Pi$ organises modes into discrete winding sectors.
      \item \textbf{Bounded-flux filtering}: the saturation bound
      $|\partial_t \chi| \leq c_\chi$ eliminates modes whose local evolution would
      exceed the substrate's relational capacity.
    \end{enumerate}
    The goal of this note is to make step~2 explicit and calculable on the simplest
    non-trivial finite graph encoding a quaternionic algebraic structure,
    as a minimal discrete proxy for $\mathrm{SU}(2)$-type sectors.

  \section{The $Q_8$ Cayley Graph}

    Let $Q_8 = \{{\pm 1, \pm i, \pm j, \pm k}\}$ denote the quaternion group of order~8.
    We define the undirected Cayley graph $\Gamma(Q_8, S)$ with generating set
    \begin{equation}
      S = \{\pm i, \pm j, \pm k\}.
    \end{equation}
    Since $S$ is closed under inversion ($s^{-1} \in S$ for all $s \in S$) and does
    not contain the identity, $\Gamma(Q_8, S)$ is a well-defined, connected, undirected,
    $6$-regular graph on $8$ vertices.

    \begin{remark}
      The choice $S = \{i, j, k\}$ (without inverses) would yield a directed graph with an
      asymmetric adjacency operator.
      For a self-adjoint Laplacian, the full symmetric set $S = \{\pm i, \pm j, \pm k\}$ is
      required.
      The resulting graph is $6$-regular, not $3$-regular.
    \end{remark}

    \subsection{Adjacency Operator and Graph Laplacian}

      The adjacency operator $A$ of $\Gamma(Q_8, S)$ acts on functions
      $f \colon Q_8 \to \mathbb{C}$ by
      \begin{equation}
        (Af)(g) = \sum_{s \in S} f(gs)
      = \sum_{s \in \{\pm i, \pm j, \pm k\}} f(gs).
      \end{equation}
      The combinatorial graph Laplacian is
      \begin{equation}
        \mathcal{L} = dI - A, \qquad d = |S| = 6,
      \end{equation}
      where $I$ is the identity on $\ell^2(Q_8)$.
      The eigenvalues of $\mathcal{L}$ are $\lambda = 6 - \mu$, where $\mu$ ranges over
      the eigenvalues of $A$.

  \section{Spectral Decomposition via Representation Theory}

    Since $Q_8$ is a finite group, $\ell^2(Q_8) \cong \mathbb{C}[Q_8]$ decomposes
    as a $Q_8$-module into irreducible representations.
    By the Peter--Weyl theorem for finite groups:
    \begin{equation}
      \mathbb{C}[Q_8] \cong \bigoplus_{\rho \in \widehat{Q_8}} V_\rho^{\oplus \dim \rho},
    \end{equation}
    where the sum runs over the five irreducible representations of $Q_8$.

    \subsection{Irreducible Representations of $Q_8$}

      $Q_8$ has five irreducible representations:
      \begin{center}
        \begin{tabular}{lcccc}
          \hline
          Label              & Dimension & $\rho(i)$ & $\rho(j)$ & $\rho(k)$ \\
          \hline
          $\rho_0$ (trivial) & 1         & $1$       & $1$       & $1$       \\
          $\rho_1$           & 1         & $1$       & $-1$      & $-1$      \\
          $\rho_2$           & 1         & $-1$      & $1$       & $-1$      \\
          $\rho_3$           & 1         & $-1$      & $-1$      & $1$       \\
          $\rho_4$ (quaternionic) & 2 & $\begin{pmatrix}
                                           i&0\\0&-i
          \end{pmatrix}$
          & $\begin{pmatrix}
               0&1\\-1&0
          \end{pmatrix}$
          & $\begin{pmatrix}
               0&i\\i&0
          \end{pmatrix}$ \\
          \hline
        \end{tabular}
      \end{center}
      The dimension check: $1^2 + 1^2 + 1^2 + 1^2 + 2^2 = 8 = |Q_8|$. \checkmark

    \subsection{Eigenvalues of the Adjacency Operator}

      For any irreducible representation $\rho$, the adjacency operator acts as the
      scalar $\mu_\rho \cdot I_{V_\rho}$ on $V_\rho$.
      Since $S$ is a union of conjugacy classes of $Q_8$, the corresponding element
      of the group algebra is central, and Schur's lemma gives
      \begin{equation}
        \mu_\rho = \frac{1}{\dim \rho} \sum_{s \in S} \chi_\rho(s),
      \end{equation}
      where $\chi_\rho$ is the character of $\rho$.

      \paragraph{1D sectors.}
        For the trivial representation $\rho_0$: $\chi_{\rho_0}(s) = 1$ for all $s \in S$,
        so $\sum_{s \in S} \chi_{\rho_0}(s) = 6$ and
      \begin{equation}
        \mu_{\rho_0} = \frac{1}{1} \cdot 6,
        \qquad
        \lambda_{\rho_0} = 6 - 6 = 0.
        \end{equation}
        For $\rho_1$: $\chi_{\rho_1}(\pm i) = 1$ and
        $\chi_{\rho_1}(\pm j) = \chi_{\rho_1}(\pm k) = -1$,
        so $\sum_{s \in S} \chi_{\rho_1}(s) = 2(1) + 4(-1) = -2$, giving
        $\mu_{\rho_1} = -2$ and $\lambda_{\rho_1} = 8$.
        By the symmetry of $Q_8$ under permutations of $\{i,j,k\}$,
        $\mu_{\rho_2} = \mu_{\rho_3} = -2$ and
        $\lambda_{\rho_2} = \lambda_{\rho_3} = 8$.

      \paragraph{2D sector.}
        For the quaternionic representation $\rho_4$, we compute the character at each
        generator.
        The matrices $\rho_4(i)$, $\rho_4(-i)$, $\rho_4(j)$, $\rho_4(-j)$, $\rho_4(k)$,
        $\rho_4(-k)$ each have trace zero: for instance,
        $\mathrm{tr}(\rho_4(i)) =
        \mathrm{tr}\bigl(\begin{smallmatrix}
                           i&0\\0&-i
        \end{smallmatrix}\bigr) = 0$,
        and $\mathrm{tr}(\rho_4(-i)) = 0$; similarly for $j$ and $k$.
        Thus $\sum_{s \in S} \chi_{\rho_4}(s) = 0$ and
      \begin{equation}
        \mu_{\rho_4} = \frac{1}{2} \cdot 0 = 0,
        \qquad
        \lambda_{\rho_4} = 6 - 0 = 6.
        \end{equation}

  \subsection{Summary of the Spectrum}

    \begin{center}
      \begin{tabular}{lccc}
        \hline
        Sector                   & $\lambda_n$ & Multiplicity & Type                       \\
        \hline
        $\rho_0$                 & $0$         & $1$          & Trivial (zero mode)        \\
        $\rho_1, \rho_2, \rho_3$ & $8$         & $3$          & Nontrivial abelian         \\
        $\rho_4$                 & $6$         & $4$          & Non-abelian (quaternionic) \\
        \hline
      \end{tabular}
    \end{center}
    Total dimension: $1 + 3 + 4 = 8$. \checkmark

  \section{Bounded-Flux Admissibility Condition}

  \subsection{Setup}

    Let the substrate field $\chi \colon Q_8 \times \mathbb{R} \to \mathbb{R}$ evolve on the
    graph, subject to the local bounded-flux constraint
    \begin{equation}
      \label{eq:flux-bound}
      |\partial_t \chi_v(t)| \leq c_\chi \qquad \forall v \in Q_8, \; \forall t.
    \end{equation}
    Consider a single-mode excitation
    \begin{equation}
      \chi_v(t) = a_n \psi_n(v) \cos(\omega_n t),
    \end{equation}
    where $\psi_n$ is an eigenmode of $\mathcal{L}$ with eigenvalue $\lambda_n$,
    normalised in $\ell^2(Q_8)$, and $a_n > 0$ is the mode amplitude.
    The time derivative at vertex $v$ is
    \begin{equation}
      \partial_t \chi_v(t) = -a_n \omega_n \psi_n(v) \sin(\omega_n t).
    \end{equation}
    The constraint~\eqref{eq:flux-bound} requires
    \begin{equation}
      |\omega_n| \cdot a_n \cdot |\psi_n(v)| \leq c_\chi
      \qquad \forall v \in Q_8.
    \end{equation}
    The binding constraint is at the vertex where $|\psi_n(v)|$ is maximal:
    \begin{equation}
      |\omega_n| \cdot a_n \cdot \|\psi_n\|_\infty \leq c_\chi.
    \end{equation}

  \subsection{Effective Amplitude and Admissibility Envelope}

    \begin{definition}
      The \emph{effective amplitude} of a mode $(\psi_n, a_n)$ is
      \begin{equation}
        A_n := a_n \|\psi_n\|_\infty.
      \end{equation}
      A mode is \emph{dynamically admissible} (in the linearised regime) if
      \begin{equation}
        \label{eq:admissibility}
        |\omega_n| \cdot A_n \leq c_\chi.
      \end{equation}
    \end{definition}

    In the linearised dispersion approximation, $\omega_n^2 \approx \lambda_n$, so the
    maximal admissible effective amplitude is
    \begin{equation}
      \label{eq:envelope}
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}.
    \end{equation}
    This defines the \emph{admissibility envelope} in the $(\lambda_n, A_n)$ plane.

    \begin{remark}
      The linearised dispersion $\omega_n^2 = \lambda_n$ is valid for
      $A_n \ll c_\chi / \sqrt{\lambda_n}$.
      The non-linear (Born-Infeld) correction modifies this relation near saturation and
      requires a modal reduction of the DBI action; it is deferred to a separate calculation.
      The envelope~\eqref{eq:envelope} should therefore be understood as the admissibility
      boundary in the weakly non-linear regime.
    \end{remark}

  \subsection{Explicit Eigenmodes in the $\rho_4$ Sector}
    \label{subsec:explicit-rho4}

    The non-abelian sector $\rho_4$ has multiplicity $4$ in $\ell^2(Q_8)$, so the
    quantity $\|\psi_n\|_\infty$ is not determined until an explicit basis of the
    eigenspace is chosen.
    A natural orthonormal basis is provided by the matrix coefficients of the
    quaternionic irreducible representation.

    For a finite group $G$ and an irreducible representation $\rho$, the standard
    Peter--Weyl basis of the corresponding isotypic component is given by
    \begin{equation}
      \psi_{uv}(g)
      =
      \sqrt{\frac{\dim \rho}{|G|}}\,\rho(g)_{uv},
      \qquad
      u,v \in \{1,\dots,\dim \rho\}.
    \end{equation}
    For $Q_8$ and $\rho_4$, this becomes
    \begin{equation}
      \psi_{uv}(g)
      =
      \frac{1}{2}\,\rho_4(g)_{uv},
      \qquad
      u,v \in \{1,2\},
    \end{equation}
    since $\dim \rho_4 = 2$ and $|Q_8| = 8$.

    Using the explicit matrices
    \begin{equation}
      \rho_4(1)=
      \begin{pmatrix}
        1 & 0 \\
        0 & 1
      \end{pmatrix},
      \qquad
      \rho_4(-1)=
      \begin{pmatrix}
        -1 & 0 \\
        0 & -1
      \end{pmatrix},
    \end{equation}
    \begin{equation}
      \rho_4(i)=
      \begin{pmatrix}
        i & 0 \\
        0 & -i
      \end{pmatrix},
      \qquad
      \rho_4(j)=
      \begin{pmatrix}
        0 & 1 \\
        -1 & 0
      \end{pmatrix},
      \qquad
      \rho_4(k)=
      \begin{pmatrix}
        0 & i \\
        i & 0
      \end{pmatrix},
    \end{equation}
    one finds that the diagonal modes $\psi_{11}, \psi_{22}$ are supported on
    $\{\pm 1, \pm i\}$, while the off-diagonal modes $\psi_{12}, \psi_{21}$ are
    supported on $\{\pm j, \pm k\}$.
    In all four cases,
    \begin{equation}
      |\psi_{uv}(g)| \in \left\{0,\frac{1}{2}\right\},
    \end{equation}
    hence
    \begin{equation}
      \|\psi_{uv}\|_\infty = \frac{1}{2},
      \qquad
      u,v \in \{1,2\}.
    \end{equation}

    For the one-dimensional nontrivial abelian sectors $\rho_1, \rho_2, \rho_3$,
    the normalised eigenmodes are the corresponding characters scaled by
    $|Q_8|^{-1/2}$.
    Since these characters take values in $\{\pm 1\}$, one has
    \begin{equation}
      \|\psi_{\rho_k}\|_\infty = \frac{1}{\sqrt{8}},
      \qquad
      k = 1,2,3.
    \end{equation}

  \subsection{Basis-Resolved Admissibility}

    \begin{proposition}[Basis-resolved admissibility on $Q_8$]
      \label{prop:basis-resolved}
      In the canonical Peter--Weyl basis of $\ell^2(Q_8)$, the maximal admissible
      effective amplitudes satisfy
      \begin{equation}
        A_{\rho_4}^{\max}
        =
        \frac{c_\chi}{\sqrt{6}}
        >
        \frac{c_\chi}{\sqrt{8}}
        =
        A_{\rho_k}^{\max},
        \qquad
        k=1,2,3,
      \end{equation}
      whereas the corresponding maximal raw amplitudes satisfy
      \begin{equation}
        a_{\rho_4}^{\max}
        =
        \frac{c_\chi}{\sqrt{6}\,\|\psi_{\rho_4}\|_\infty}
        =
        \frac{2}{\sqrt{6}}\,c_\chi
        \approx 0.816\,c_\chi,
      \end{equation}
      \begin{equation}
        a_{\rho_k}^{\max}
        =
        \frac{c_\chi}{\sqrt{8}\,\|\psi_{\rho_k}\|_\infty}
        =
        c_\chi,
        \qquad
        k=1,2,3.
      \end{equation}
      Thus the non-abelian sector is favoured in effective amplitude, but not in raw
      amplitude.
    \end{proposition}

    \begin{proof}
      The effective-amplitude statement follows from
      $A_n^{\max} = c_\chi/\sqrt{\lambda_n}$ together with
      $\lambda_{\rho_4} = 6$ and $\lambda_{\rho_k} = 8$.
      The raw-amplitude statement follows from
      $a_n^{\max} = A_n^{\max}/\|\psi_n\|_\infty$, using
      $\|\psi_{\rho_4}\|_\infty = 1/2$ and
      $\|\psi_{\rho_k}\|_\infty = 1/\sqrt{8}$.
    \end{proof}

    \begin{remark}[Interpretation]
        The bounded-flux constraint acts on the local quantity
        $|\omega_n|\,a_n\,|\psi_n(v)|$, and therefore selects modes through the effective
        amplitude $A_n = a_n \|\psi_n\|_\infty$, not through the raw amplitude $a_n$ alone.
        The non-abelian advantage on $Q_8$ is therefore not a larger absolute amplitude
        capacity, but a larger admissible effective window per unit spectral weight.
        Equivalently, the $\rho_4$ sector is less constrained by bounded-flux filtering
        in the physically relevant local variable.
    \end{remark}

  \subsection{Weakly Non-linear Modal DBI Reduction}
    \label{subsec:weakly-nonlinear-dbi}

    To incorporate the first Born--Infeld correction, consider the discrete DBI
    Lagrangian
    \begin{equation}
      \label{eq:dbi-lagrangian}
      L
      =
      -c_\chi^2 \sum_v \sqrt{1 - \dot{\chi}_v^2/c_\chi^2}
      - \tfrac{1}{2} \sum_{v,w} \chi_v \mathcal{L}_{vw} \chi_w.
    \end{equation}
    For a single eigenmode,
    \begin{equation}
      \chi_v(t) = q(t)\psi_n(v),
    \end{equation}
    with $\mathcal{L}\psi_n = \lambda_n \psi_n$ and $\|\psi_n\|_2 = 1$, this reduces to
    the exact one-mode Lagrangian
    \begin{equation}
      \label{eq:exact-mode-lagrangian}
      L_n[q]
      =
      -c_\chi^2 \sum_v \sqrt{1 - \psi_n(v)^2 \dot{q}^{\,2}/c_\chi^2}
      - \tfrac{1}{2}\lambda_n q^2.
    \end{equation}

    Expanding the square root for $|\dot{q}| \ll c_\chi$ gives
    \begin{equation}
      \label{eq:mode-lagrangian-expansion}
      L_n[q]
      =
      \mathrm{const}
      + \tfrac{1}{2}\dot{q}^{\,2}
      + \frac{\kappa_n}{8c_\chi^2}\dot{q}^{\,4}
      - \tfrac{1}{2}\lambda_n q^2
      + \mathcal{O}(c_\chi^{-4}),
    \end{equation}
    where
    \begin{equation}
      \label{eq:kappa-def}
      \kappa_n := \sum_v \psi_n(v)^4.
    \end{equation}
    The corresponding Euler--Lagrange equation is
    \begin{equation}
      \label{eq:weakly-nonlinear-mode-equation}
      \left(
        1 + \frac{3\kappa_n}{2c_\chi^2}\dot{q}^{\,2}
      \right)\ddot{q}
      + \lambda_n q
      = 0.
    \end{equation}

    Using the harmonic-balance ansatz
    \begin{equation}
      q(t) \approx a_n \cos(\omega_n t),
    \end{equation}
    and keeping only the fundamental harmonic, one obtains the first
    amplitude-dependent correction
    \begin{equation}
      \label{eq:omega-shift-square}
      \omega_n^2(a_n)
      \approx
      \lambda_n
      \left(
        1 - \frac{3\kappa_n\lambda_n a_n^2}{8c_\chi^2}
      \right),
    \end{equation}
    or equivalently
    \begin{equation}
      \label{eq:omega-shift}
      \omega_n(a_n)
      \approx
      \sqrt{\lambda_n}
      \left(
        1 - \frac{3\kappa_n\lambda_n a_n^2}{16c_\chi^2}
      \right).
    \end{equation}
    The DBI correction is therefore a softening non-linearity: the frequency
    decreases as the mode amplitude grows.

    In the canonical $Q_8$ basis, the quartic shape factor is explicit.
    For the one-dimensional nontrivial abelian sectors,
    \begin{equation}
      \label{eq:kappa-abelian}
      \kappa_{\rho_k}
      =
      \sum_{v \in Q_8} \left(\frac{1}{\sqrt{8}}\right)^4
      =
      \frac{1}{8},
      \qquad
      k = 1,2,3.
    \end{equation}
    For each canonical basis vector in the $\rho_4$ sector,
    \begin{equation}
      \label{eq:kappa-nonabelian}
      \kappa_{\rho_4}
      =
      4 \left(\frac{1}{2}\right)^4
      =
      \frac{1}{4}.
    \end{equation}
    Hence the raw-amplitude correction is stronger in the non-abelian sector,
    reflecting the larger localisation of the corresponding basis vectors.

    However, rewriting the correction in terms of the effective amplitude
    $A_n = a_n\|\psi_n\|_\infty$ gives
    \begin{equation}
      \label{eq:omega-shift-effective}
      \omega_n^2(A_n)
      \approx
      \lambda_n
      \left(
        1
        - \frac{3\lambda_n A_n^2}{8c_\chi^2}
        \frac{\kappa_n}{\|\psi_n\|_\infty^2}
      \right).
    \end{equation}
    In the canonical $Q_8$ basis,
    \begin{equation}
      \frac{\kappa_{\rho_k}}{\|\psi_{\rho_k}\|_\infty^2}
      =
      \frac{(1/8)}{(1/\sqrt{8})^2}
      =
      1,
      \qquad
      \frac{\kappa_{\rho_4}}{\|\psi_{\rho_4}\|_\infty^2}
      =
      \frac{(1/4)}{(1/2)^2}
      =
      1.
    \end{equation}
    Therefore
    \begin{equation}
      \label{eq:omega-shift-effective-q8}
      \omega_n^2(A_n)
      \approx
      \lambda_n
      \left(
        1 - \frac{3\lambda_n A_n^2}{8c_\chi^2}
      \right)
    \end{equation}
    in both the nontrivial abelian and non-abelian sectors.
    At fixed effective amplitude, the non-abelian sector thus remains less penalised
    by the weakly non-linear DBI correction, solely because $\lambda_{\rho_4}=6$
    is smaller than $\lambda_{\rho_k}=8$.

  \subsection{Non-perturbative DBI envelope}
    \label{subsec:nonperturbative-dbi}

    The weakly non-linear reduction derived in
    Section~\ref{subsec:weakly-nonlinear-dbi} provides the leading
    amplitude correction to the dispersion relation.
    However, the discrete Dirac--Born--Infeld dynamics allows the
    single-mode problem to be solved exactly on the $Q_8$ Cayley graph.

    Consider again a single eigenmode
    \[
      \chi_v(t) = q(t)\psi_n(v),
    \]
    with $\mathcal{L}\psi_n = \lambda_n \psi_n$ and $\|\psi_n\|_2=1$.
    Substituting this ansatz into the DBI Lagrangian
    Eq.~\eqref{eq:dbi-lagrangian} gives the exact one-mode Lagrangian
    \begin{equation}
      L_n[q]
      =
      -c_\chi^2
      \sum_{v\in Q_8}
      \sqrt{1-\frac{\psi_n(v)^2\dot q^{\,2}}{c_\chi^2}}
      -
      \frac12\lambda_n q^2 .
    \end{equation}

    Because the eigenmodes on the $Q_8$ Cayley graph have a simple
    support structure, the DBI sum can be evaluated explicitly.

    \subsubsection*{Abelian sectors}

      For the nontrivial abelian representations $\rho_1,\rho_2,\rho_3$,
      the normalized eigenmodes satisfy
      \[
        |\psi_n(v)| = \frac{1}{\sqrt{8}}
        \qquad
        \forall v\in Q_8 .
      \]

      The Lagrangian therefore becomes
      \begin{equation}
        L
        =
        -8c_\chi^2
        \sqrt{1-\frac{\dot q^{\,2}}{8c_\chi^2}}
        -
        \frac12\lambda_n q^2 .
      \end{equation}

    \subsubsection*{Non-abelian sector}

      For the quaternionic sector $\rho_4$ in the canonical basis,
      the eigenmodes satisfy
      \[
        |\psi_n(v)| =
        \begin{cases}
          \frac12 & \text{on four vertices}, \\
          0 & \text{on four vertices}.
        \end{cases}
      \]

      The DBI sum therefore reduces to
      \begin{equation}
        L
        =
        -4c_\chi^2
        \sqrt{1-\frac{\dot q^{\,2}}{4c_\chi^2}}
        -4c_\chi^2
        -
        \frac12\lambda_n q^2 .
      \end{equation}

      The constant term does not affect the dynamics, and the effective
      Lagrangian becomes
      \begin{equation}
        L
        =
        -4c_\chi^2
        \sqrt{1-\frac{\dot q^{\,2}}{4c_\chi^2}}
        -
        \frac12\lambda_n q^2 .
      \end{equation}

    \subsubsection*{Energy conservation}

      Since the Lagrangian is time independent, the corresponding Hamiltonian
      is conserved. One obtains
      \begin{equation}
        E
        =
        4c_\chi^2
        \frac{1}{\sqrt{1-\frac{\dot q^{\,2}}{4c_\chi^2}}}
        +
        \frac12\lambda_n q^2 .
      \end{equation}

      At the turning points $\dot q=0$, where $q=\pm a_n$, the energy becomes
      \begin{equation}
        E
        =
        4c_\chi^2
        +
        \frac12\lambda_n a_n^2 ,
      \end{equation}
      where $a_n$ denotes the raw oscillation amplitude of the modal coordinate.

    \subsubsection*{Velocity bound}

      The DBI square root imposes the kinematic bound
      \begin{equation}
        |\dot q| < 2c_\chi .
      \end{equation}

      Expressed in terms of the field variable,
      \[
        \chi_v = q\psi_n(v),
      \]
      this corresponds exactly to the local flux constraint
      \[
        |\partial_t \chi_v| \le c_\chi .
      \]

    \subsubsection*{Amplitude bound}

      The turning point occurs when the maximal vertex saturates the flux bound,
      \[
        |\partial_t\chi_v|
        =
        |\dot q|\,|\psi_n(v)|
        =
        c_\chi .
      \]

      Since $\|\psi_n\|_\infty$ is the maximal value of $|\psi_n(v)|$, this gives
      \begin{equation}
        a_n^{\max}
        =
        \frac{c_\chi}{\|\psi_n\|_\infty\sqrt{\lambda_n}} .
      \end{equation}

      The corresponding effective amplitude
      \[
        A_n = a_n\|\psi_n\|_\infty
      \]
      therefore satisfies
      \begin{equation}
        A_n^{\max}
        =
        \frac{c_\chi}{\sqrt{\lambda_n}} .
      \end{equation}

    \subsubsection*{Exact admissibility envelope}

      We thus obtain the non-perturbative result

      \begin{equation}
        A_n^{\max}
        =
        \frac{c_\chi}{\sqrt{\lambda_n}},
      \end{equation}

      which coincides exactly with the admissibility envelope derived from the
      linear bounded-flux argument.

      The weakly non-linear correction therefore does not modify the
      location of the admissibility boundary.
      Instead, the DBI dynamics smoothly interpolates between the linear regime
      and a velocity-saturated regime as the amplitude approaches the bound.

      This shows that the spectral admissibility envelope identified in
      Step~1 is not merely a linear approximation but the exact saturation
      boundary for the single-mode DBI dynamics on the $Q_8$ Cayley graph.

  \section{Main Result}

  \begin{proposition}[Spectral admissibility hierarchy on $Q_8$]
    \label{prop:main}
    In the undirected Cayley graph $\Gamma(Q_8, \{\pm i, \pm j, \pm k\})$,
    under the local bounded-flux constraint~\eqref{eq:flux-bound} and linearised
    dispersion, the maximal admissible effective amplitude satisfies
    \begin{equation}
      A^{\max}_{\rho_4} = \frac{c_\chi}{\sqrt{6}} > \frac{c_\chi}{\sqrt{8}}
      = A^{\max}_{\rho_k}, \quad k = 1, 2, 3.
    \end{equation}
    The non-abelian sector $\rho_4$ occupies a strictly larger admissible region in the
    space of effective mode amplitudes than any nontrivial abelian sector, by a factor
    \begin{equation}
      \frac{A^{\max}_{\rho_4}}{A^{\max}_{\rho_k}}
      =
      \sqrt{\frac{8}{6}}
      =
      \sqrt{\frac{4}{3}}
      \approx 1.155.
    \end{equation}
  \end{proposition}

  \begin{proof}
    Direct substitution of $\lambda_{\rho_4} = 6$ and $\lambda_{\rho_k} = 8$
    into the admissibility envelope~\eqref{eq:envelope}.
  \end{proof}

  \begin{remark}
    The trivial sector $\rho_0$ has $\lambda = 0$ and thus formally unbounded $A^{\max}$.
    It corresponds to the uniform zero mode of the connected graph.
    In the present analysis it is factored out, since we are interested only in
    non-trivial excitations above the connected background.
  \end{remark}

  \section{Interpretation}

  Proposition~\ref{prop:main}, together with the basis-resolved
  Proposition~\ref{prop:basis-resolved}, establishes that, at fixed substrate capacity
  $c_\chi$, the non-abelian irreducible sector of the quaternion Cayley graph sustains
  a larger local effective amplitude before violating the bounded-flux constraint,
  even though its raw amplitude bound is lower than that of the nontrivial abelian
  sectors.

  In the language of the Cosmochrony programme:

  \begin{itemize}
    \item The admissibility envelope $A_n^{\max} = c_\chi / \sqrt{\lambda_n}$ defines a
    \emph{geometry of exclusion} in the space of relational modes: modes above the
    envelope are dynamically inadmissible.
    \item The non-abelian sector $\rho_4$ lies strictly below this envelope for a larger
    range of effective amplitudes than the nontrivial abelian sectors
    $\rho_1, \rho_2, \rho_3$: the bounded-flux constraint is less restrictive for
    non-abelian modes, enlarging their admissible window by a factor of $\sqrt{4/3}$.
    \item This advantage is a statement about the physically relevant local variable
    $A_n = a_n \|\psi_n\|_\infty$, not about the raw amplitude $a_n$.
    In the canonical basis, the $\rho_4$ modes are more localised
    ($\|\psi\|_\infty = 1/2$) than the abelian ones
    ($\|\psi\|_\infty = 1/\sqrt{8}$), so the non-abelian sector is more constrained
    in raw amplitude even while remaining less constrained in effective amplitude.
    \item The weakly non-linear DBI correction is governed by the quartic shape factor
    $\kappa_n = \sum_v \psi_n(v)^4$.
    In raw amplitude, the canonical $\rho_4$ modes are more strongly penalised than
    the abelian ones because they are more localised.
    In effective amplitude, this localisation factor cancels on $Q_8$, so the first
    DBI correction again favours the non-abelian sector through its smaller
    eigenvalue.
    \item This provides a minimal admissibility advantage for non-abelian relational
    modes, both at the linear level and at the first weakly non-linear DBI order,
    without any dynamical postulate beyond the local flux bound.
  \end{itemize}

  \begin{remark}[Real field and complex eigenmodes]
      The substrate field $\chi$ is taken real-valued.
      The natural eigenmodes of the 2D sector are matrix coefficients of $\rho_4$,
      which are complex-valued.
      One may either work in $\ell^2(Q_8, \mathbb{C})$ and take real and imaginary parts to
      obtain real eigenmodes, or directly choose a real basis for each eigenspace.
      The local admissibility criterion is intrinsically defined, but its expression in
      terms of $(a_n,\psi_n)$ depends on the chosen basis within a degenerate eigenspace
      through $\|\psi_n\|_\infty$.
      This basis dependence is resolved explicitly in
      Section~\ref{subsec:explicit-rho4}.
  \end{remark}

  This result does not yet establish:
  \begin{itemize}
    \item that the non-abelian sector will dominate in a full dynamical simulation;
    \item the origin of the eigenvalue gap $\lambda_{\rho_4} = 6 < 8 = \lambda_{\rho_k}$
    from the topological structure of $\Pi$ alone;
    \item the fully non-linear Born--Infeld correction to the admissibility envelope
    near saturation beyond the weakly non-linear regime.
  \end{itemize}
  These remain open and constitute the next steps of the programme.

  \section{Synthesis: Spectral Filtering by Bounded Flux}

  The analysis performed in this note identifies the effective amplitude
  \[
    A_n = a_n \|\psi_n\|_\infty
  \]
  as the physically relevant variable controlling the saturation of the
  substrate.

  \subsection{Local Flux Constraint as a Spectral Filter}

    The bounded-flux condition
    \[
      |\partial_t \chi_v| \le c_\chi
    \]
    acts locally on each vertex of the relational substrate.

    For a single eigenmode
    \[
      \chi_v(t) = a_n \psi_n(v) \cos(\omega_n t),
    \]
    the constraint becomes
    \[
      |\omega_n|\,a_n\,\|\psi_n\|_\infty \le c_\chi.
    \]

    Introducing the effective amplitude
    \[
      A_n = a_n \|\psi_n\|_\infty
    \]
    yields the admissibility envelope
    \[
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}.
    \]

    A key consequence is that the localization of the mode does not alter the
    admissibility bound itself.
    Localization only determines how the raw amplitude $a_n$ translates into
    a local flux impact on the substrate.
    The selection mechanism is therefore governed purely by the spectral
    eigenvalue $\lambda_n$.

  \subsection{DBI Softening as Dynamical Spectral Filtering}

    In the weakly non-linear regime the discrete Dirac--Born--Infeld dynamics
    produces an amplitude-dependent frequency

    \[
      \omega_n^2(A_n)
      \approx
      \lambda_n
      \left(
        1 -
        \frac{3\lambda_n A_n^2}{8c_\chi^2}
      \right).
    \]

    This simplified expression follows from the identity

    \[
      \frac{\kappa_n}{\|\psi_n\|_\infty^2}=1
    \]

    which holds for all non-trivial sectors of the $Q_8$ Cayley graph.

    The non-linear correction therefore depends only on the Laplacian
    eigenvalue.
    Modes with larger $\lambda_n$ undergo stronger softening as the
    effective amplitude increases.

    The relational substrate thus behaves as a dynamical low-pass spectral
    filter.

  \subsection{Conceptual Interpretation}

    The mechanism uncovered in this toy model can be summarized as

    \[
      \text{sub\-strate capacity}
      \;\rightarrow\;
      \text{spectral admissibility}
      \;\rightarrow\;
      \text{nonlinear filtering}
      \;\rightarrow\;
      \text{surviving sectors}.
    \]

    On the $Q_8$ Cayley graph the smallest non-trivial eigenvalue

    \[
      \lambda_{\rho_4}=6
    \]

    belongs to the quaternionic non-abelian sector.
    This sector therefore possesses the largest admissible amplitude window
    and is the least affected by DBI softening.

    In this sense, physical structures correspond to the spectral
    components that remain dynamically admissible under the bounded
    capacity of the relational substrate.

    Matter may thus be interpreted as the dynamical residue of the spectral
    filtering imposed by the substrate.

    The non-perturbative DBI reduction derived in
    Section~\ref{subsec:nonperturbative-dbi} shows that the admissibility
    envelope

    \[
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}
    \]

    is not merely a linear approximation.
    It coincides with the exact saturation boundary of the one-mode
    DBI dynamics on the $Q_8$ Cayley graph.

    As the effective amplitude approaches this bound,

    \[
      A_n \rightarrow A_n^{\max},
    \]

    the modal dynamics enters a velocity-saturated regime in which the
    local flux constraint
    \[
      |\partial_t \chi_v| \le c_\chi
    \]
    is approached at the maximally excited vertex.

    In this regime the oscillation no longer grows in amplitude,
    and the dynamics becomes increasingly constrained by the
    substrate capacity.
    Energy injected beyond the admissible region cannot be sustained
    within the corresponding spectral component and is therefore
    redistributed toward modes with smaller spectral weight.

    The bounded relational capacity of the substrate therefore acts
    not merely as a spectral cut-off but as a dynamical selector,
    favouring the most resilient spectral sectors.

  \subsection{Beyond the $Q_8$ Toy Model}

    The identity
    \[
      \frac{\kappa_n}{\|\psi_n\|_\infty^2}=1
    \]
    is a special feature of the small and highly symmetric $Q_8$ graph.

    On larger graphs this ratio will generally depend on the structure of
    the eigenmodes.
    The non-linear DBI correction may therefore amplify the spectral
    selection mechanism rather than acting uniformly across sectors.

    This motivates extending the analysis to larger Cayley graphs,
    including Ramanujan families such as the Lubotzky--Phillips--Sarnak
    constructions on $\mathrm{SL}(2,\mathbb{F}_p)$.

  \subsection{Next Step: Dynamical Survival}

    The next stage of the programme is to test this mechanism dynamically.

    The proposed protocol is to initialise the substrate with a mixture of
    spectral modes close to the admissibility envelope and evolve the
    system under the discrete DBI dynamics.

    If spectral cascades preferentially accumulate in the lowest
    admissible sector, this would demonstrate a genuine dynamical
    selection of symmetry sectors driven by the bounded relational
    capacity of the substrate.

    Sections~2–7 established the spectral admissibility hierarchy on the $Q_8$ Cayley graph.
    We now identify the structural origin of this hierarchy and its
    possible connection with the $\mathrm{SU}(2)$ geometry underlying spin.

    % ============================================================
% Spectral Origin of Non-Abelian Resilience
%              and the SU(2) Emergence Thread
% ============================================================
% This section extends the completed Steps 1–3 of the present note.
% It provides the algebraic explanation of the ratio sqrt(4/3),
% identifies the connection with the S^3 / SU(2) spectral geometry,
% and outlines the programme toward SU(2) emergence via bounded flux.

  \section{Structural Origin of the Admissibility Hierarchy and the SU(2) Emergence Thread}
  \label{sec:step4}

  \subsection{The ratio $\sqrt{4/3}$ is not accidental}
    \label{sec:step4-origin}

    The admissibility hierarchy established in Proposition~2,
    \begin{equation}
      \frac{A^{\max}_{\rho_4}}{A^{\max}_{\rho_k}}
      = \sqrt{\frac{8}{6}} = \sqrt{\frac{4}{3}} \approx 1.155,
      \label{eq:ratio}
    \end{equation}
    has a precise internal origin in the $Q_8$ Cayley spectrum.
    It originates entirely from the Laplacian eigenvalues $\lambda_{\rho_4} = 6$ and
    $\lambda_{\rho_k} = 8$ ($k = 1,2,3$).
    The ratio $8/6$ is in turn fixed by the adjacency eigenvalues $\mu_{\rho_k} = -2$
    and $\mu_{\rho_4} = 0$ through $\lambda = d - \mu$ with $d = 6$.

    The central observation is that $\mu_{\rho_4} = 0$: the quaternionic sector
    lies \emph{exactly at the centre} of the adjacency spectrum
    $\mu \in [-d, d] = [-6, 6]$.
    This is a structural consequence of the character theory of $Q_8$,
    not a numerical coincidence.

  \subsection{Why $\mu_{\rho_4} = 0$: character-theoretic argument}
    \label{sec:step4-character}

    For any Cayley graph on a finite group $G$ with symmetric generating set $S$,
    Schur's lemma gives the adjacency eigenvalue in the irreducible sector $\rho$ as
    \begin{equation}
      \mu_\rho = \frac{1}{\dim \rho} \sum_{s \in S} \chi_\rho(s).
      \label{eq:mu-character}
    \end{equation}
    For the quaternionic representation $\rho_4$ of $Q_8$, the generating set
    $S = \{\pm i, \pm j, \pm k\}$ (with $|S| = 6$, fixing the regularity degree
    $d = 6$) maps to the matrices
    \begin{equation}
      \rho_4({\pm i}) = \pm\begin{pmatrix}
                             i & 0 \\ 0 & -i
      \end{pmatrix},
      \quad
      \rho_4({\pm j}) = \pm\begin{pmatrix}
                             0 & 1 \\ -1 & 0
      \end{pmatrix},
      \quad
      \rho_4({\pm k}) = \pm\begin{pmatrix}
                             0 & i \\ i & 0
      \end{pmatrix}.
    \end{equation}
    All six generator matrices are traceless: they are proportional to the Pauli
    matrices $\sigma_x, \sigma_y, \sigma_z$, which satisfy
    $\mathrm{tr}(\sigma_\alpha) = 0$.
    Therefore
    \begin{equation}
      \chi_{\rho_4}(s) = \mathrm{tr}(\rho_4(s)) = 0
      \qquad \forall\, s \in S,
      \label{eq:vanishing-character}
    \end{equation}
    and the character sum in Eq.~\eqref{eq:mu-character} vanishes identically,
    giving $\mu_{\rho_4} = 0$.

    This is a structural property: the generators of $Q_8$ map to traceless matrices
    under $\rho_4$ because $Q_8$ is generated by the unit quaternions $i, j, k$,
    whose matrix images are exactly the imaginary units of $\mathrm{SU}(2)$.
    The tracelessness is not a special feature of the graph but of the embedding
    $Q_8 \hookrightarrow \mathrm{SU}(2)$.

    \begin{quote}
      The ratio $\sqrt{4/3}$ appearing in the admissibility hierarchy is not an
      accidental numerical feature of the toy model.
      It originates from the character structure of the quaternionic
      representation of $Q_8$, for which the adjacency eigenvalue satisfies
      $\mu_{\rho_4} = 0$.
      The non-abelian sector occupies the central position of the adjacency
      spectrum, resulting in the smallest non-trivial Laplacian eigenvalue and
      hence the largest admissible amplitude window.
    \end{quote}

  \subsection{$Q_8$ as the minimal discrete encoding of spin}
    \label{sec:step4-spin}

    The connection between $Q_8$ and spin is not coincidental but topological.
    The rotation group $SO(3)$ has fundamental group
    $\pi_1(SO(3)) = \mathbb{Z}_2$,
    meaning that a rotation by $2\pi$ cannot be continuously contracted to the
    identity: one must rotate by $4\pi$ to return topologically to the
    starting point.
    This is the origin of half-integer spin.

    The simply-connected double cover of $SO(3)$ is $\mathrm{SU}(2)$,
    which is topologically the three-sphere $S^3$:
    \begin{equation}
      \mathrm{SU}(2) \cong \big\{a_0 I + i a_1 \sigma_x + i a_2 \sigma_y
      + i a_3 \sigma_z \;:\; a_0^2 + a_1^2 + a_2^2 + a_3^2 = 1\big\}
      \cong S^3.
      \label{eq:SU2-S3}
    \end{equation}
    The unit quaternions $\{q = a_0 + a_1 i + a_2 j + a_3 k : |q| = 1\}$ are
    in bijection with $\mathrm{SU}(2)$ via the identification
    $i \leftrightarrow i\sigma_x$, $j \leftrightarrow i\sigma_y$,
    $k \leftrightarrow i\sigma_z$.

    The group $Q_8 = \{\pm 1, \pm i, \pm j, \pm k\}$ is the smallest
    non-abelian subgroup of the unit quaternions, hence the smallest
    non-abelian subgroup of $\mathrm{SU}(2)$.
    Its elements have the following interpretation as spin rotations:
    $-1$ is a $2\pi$-rotation (the spinorial sign flip);
    $\pm i$, $\pm j$, $\pm k$ are $\pi$-rotations around the three coordinate
    axes.
    $Q_8$ therefore captures the minimal non-abelian subgroup structure
    compatible with the $\mathrm{SU}(2)$ spin algebra:
    it contains both the $2\pi$ phase ($-1$) and the three independent
    rotation axes ($\pm i$, $\pm j$, $\pm k$).

    The matrices $i\sigma_x$, $i\sigma_y$, and $i\sigma_z$ generate the Lie algebra
    $\mathfrak{su}(2)$ under commutation, so
    \begin{equation}
      Q_8
      \xrightarrow{\;\rho_4\;}
      \text{Pauli algebra}
      \;\longrightarrow\;
      \mathfrak{su}(2)
      \;\longrightarrow\;
      \mathrm{SU}(2)
      \;\longrightarrow\;
      \text{spin}.
    \end{equation}
    The spectral result of this note---that bounded relational flux favours
    the $\rho_4$ sector---therefore says that bounded substrate capacity
    preferentially sustains precisely those modes encoding the spin structure
    of $\mathrm{SU}(2)$.

  \subsection{Connection with the $S^3$ spectral invariant $\lambda_2/\lambda_1 = 8/3$}
    \label{sec:step4-S3}

    On the three-sphere $S^3$, the Laplacian eigenvalues are
    $\lambda_\ell = \ell(\ell + 2)$ with multiplicity $(\ell+1)^2$, giving
    \begin{equation}
      \ell = 0:\;\lambda_0 = 0, \qquad
      \ell = 1:\;\lambda_1 = 3, \qquad
      \ell = 2:\;\lambda_2 = 8, \qquad
      \ell = 3:\;\lambda_3 = 15.
    \end{equation}
    The fundamental spectral ratio of $S^3$ is
    \begin{equation}
      \frac{\lambda_2}{\lambda_1}\bigg|_{S^3} = \frac{8}{3}.
      \label{eq:S3-ratio}
    \end{equation}
    This ratio is the primary spectral invariant used elsewhere in the
    Cosmochrony programme to characterise the $S^3$ geometry from the
    relational spectrum~\cite{Spectral121}.

    The ratio from the present $Q_8$ calculation,
    \begin{equation}
      \frac{\lambda_{\rho_k}}{\lambda_{\rho_4}} = \frac{8}{6} = \frac{4}{3},
    \end{equation}
    is related to the $S^3$ ratio by
    \begin{equation}
      \frac{\lambda_{\rho_k}}{\lambda_{\rho_4}}
      = \frac{1}{2} \cdot \frac{\lambda_2}{\lambda_1}\bigg|_{S^3}.
      \label{eq:ratio-relation}
    \end{equation}
    This numerical relation suggests a possible connection between the coarse
    spectral structure of the $Q_8$ Cayley graph and the continuum spectrum
    of $S^3$.
    At present this relation should be regarded as heuristic rather than as a
    derived theorem.
    The factor $1/2$ may reflect the extreme coarseness of the $Q_8$
    approximation of $S^3$, which has only 8 vertices.
    Testing whether this factor approaches unity for larger subgroups such as
    $2I$ is an open question.

    Two independent computations in the Cosmochrony programme nonetheless
    point toward the same $S^3$ geometry: the spectral reconstruction ratio
    $8/3$ from~\cite{Spectral121}, and the admissibility ratio $4/3$ derived
    here.

  \subsection{The SU(2) emergence programme}
    \label{sec:step4-programme}

    The $Q_8$ result motivates a systematic programme toward $\mathrm{SU}(2)$
    emergence via spectral admissibility.
    The key chain of inclusions is
    \begin{equation}
      Q_8 \;\subset\; 2I \;\subset\; \mathrm{SU}(2),
      \label{eq:inclusion-chain}
    \end{equation}
    where $2I$ is the binary icosahedral group of order 120, the
    \emph{largest} finite subgroup of $\mathrm{SU}(2)$, which is a double cover
    of $A_5 \cong \mathrm{PSL}(2, \mathbb{F}_5)$.

    The irreducible representations of $2I$ have dimensions
    $1, 2, 2', 3, 3', 4, 4', 5, 6$,
    which reproduce the low-spin content of $\mathrm{SU}(2)$:
    spin-$1/2$ (dim~2), spin-$1$ (dim~3), spin-$3/2$ (dim~4),
    spin-$2$ (dim~5), spin-$5/2$ (dim~6).
    The admissibility envelope $A^{\max}_n = c_\chi/\sqrt{\lambda_n}$ applied
    to Cayley graphs of $2I$ should therefore produce a \emph{low-spin survival}
    hierarchy, in which lower-spin representations---carrying smaller Laplacian
    eigenvalues---are favoured.

    \begin{remark}[Heuristic status of low-spin survival]
  The statement that lower-spin representations correspond to smaller
  Laplacian eigenvalues on $\mathrm{Cay}(2I, S)$ is a structural hypothesis,
  not a theorem.
  It holds for $Q_8$ and is consistent with the general pattern that
  lower-dimensional representations tend to sit lower in the Cayley graph
  spectrum for groups approximating compact Lie groups, but it requires
  explicit verification on $2I$.
    \end{remark}

    The asymptotic step is provided by the Ramanujan graphs of
    Lubotzky, Phillips, and Sarnak~\cite{LPS1988}, which are Cayley graphs of
    $\mathrm{SL}(2, \mathbb{F}_p)$ designed to approximate $S^3$ with
    near-optimal spectral gap.
    These graphs have the property that all non-trivial eigenvalues satisfy
    $|\mu| \leq 2\sqrt{d-1}$ (the Alon--Boppana bound), making them
    optimal expanders.
    The bounded-flux admissibility mechanism studied here may therefore be
    connected to expander graph theory via the spectral gap.

    The full programme is:
    \begin{equation}
      Q_8
      \;\xrightarrow{\;\text{Step 4}\;}\;
      2I
      \;\xrightarrow{\;\text{Step 5}\;}\;
      \mathrm{SL}(2, \mathbb{F}_p)
      \;\xrightarrow{\;p \to \infty\;}\;
      \mathrm{SU}(2).
    \end{equation}
    At each stage the prediction is:
    \begin{equation}
      \text{bounded flux}
      \;\Longrightarrow\;
      \text{spectral hierarchy}
      \;\Longrightarrow\;
      \text{low-spin survival}
      \;\Longrightarrow\;
      \text{SU(2) emergence}.
    \end{equation}

  \subsection{Physical interpretation}
    \label{sec:step4-physics}

    The mechanism uncovered in this toy model has a precise physical reading
    within the Cosmochrony framework.

    The non-abelian quaternionic sector has zero net correlation with the
    generators of the graph ($\mu_{\rho_4} = 0$): it neither oscillates
    \emph{with} the graph structure nor against it.
    It is spectrally neutral with respect to the dynamics.
    This neutrality is the source of its resilience: the bounded-flux
    constraint imposes less penalty on modes that do not couple to the
    relational connectivity.

    By contrast, the abelian sectors have $\mu_{\rho_k} = -2$, meaning that they
    are anti-correlated with the generators.
    They therefore pay a higher spectral price under bounded-flux filtering.

    The mechanism is therefore not gauge symmetry, not dynamical postulate,
    and not fine-tuning.
    It is a purely structural consequence of the embedding
    $Q_8 \hookrightarrow \mathrm{SU}(2)$ and the bounded capacity of the
    relational substrate:
    \begin{equation}
      \text{bounded relational capacity}
      \;\Longrightarrow\;
      \text{non-abelian sector resilience}.
    \end{equation}

    If this mechanism persists along the chain~\eqref{eq:inclusion-chain},
    the gauge symmetry group $\mathrm{SU}(2)$ would not need to be postulated.
    It would emerge as the limiting residue of spectral filtering under
    bounded substrate capacity.

    \bigskip
    \noindent\textbf{Summary of Step~4.}
    The $Q_8$ admissibility hierarchy is explained by the vanishing character
    sum $\sum_{s \in S} \chi_{\rho_4}(s) = 0$, itself a consequence of the
    tracelessness of the Pauli matrices.
    The resulting ratio $\sqrt{4/3}$ is numerically related by a factor $1/2$
    to the fundamental spectral ratio $8/3$ of $S^3$; this relation is
    heuristic at present but both quantities point toward the same
    $\mathrm{SU}(2) \cong S^3$ geometry.
    The $Q_8$ result is therefore not merely a toy model curiosity: it is
    the first discrete instance of a potential mechanism by which bounded
    relational capacity selects $\mathrm{SU}(2)$-type symmetry sectors.

  \section{Next Steps}
  \label{sec:next-steps}

  The present note completes the first two stages of the spectral admissibility
  programme on the $Q_8$ Cayley graph.

  \paragraph{Step 1: spectral admissibility.}
    The bounded-flux constraint
    $|\partial_t \chi_v| \le c_\chi$
    selects admissible modes through the effective amplitude
    $A_n = a_n \|\psi_n\|_\infty$,
    leading to the envelope
    \[
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}.
    \]
    On $Q_8$, this produces a strict hierarchy in which the quaternionic
    non-abelian sector $\rho_4$ admits the largest effective amplitude window.

  \paragraph{Step 2: weakly non-linear DBI reduction.}
    The discrete Dirac--Born--Infeld dynamics has been reduced to a single-mode
    effective equation, yielding the amplitude-dependent frequency
    \[
      \omega_n^2(A_n)
      \approx
      \lambda_n
      \left(
        1 - \frac{3\lambda_n A_n^2}{8c_\chi^2}
      \right).
    \]
    On the $Q_8$ Cayley graph the ratio
    $\kappa_n/\|\psi_n\|_\infty^2 = 1$
    for all non-trivial sectors, implying that the non-linear softening depends
    only on the Laplacian eigenvalue.
    The spectral hierarchy obtained in Step~1 therefore remains stable under the
    leading DBI correction.

  \paragraph{Step 3: non-perturbative DBI envelope.}
    The next task is to move beyond the perturbative expansion in $c_\chi^{-2}$
    and derive the full amplitude-dependent period from the exact one-mode
    Lagrangian~\eqref{eq:exact-mode-lagrangian}.
    This can be obtained either through elliptic-integral methods or direct
    numerical quadrature.
    The resulting period--amplitude relation will determine the exact
    non-perturbative admissibility envelope replacing the linear boundary
    \eqref{eq:envelope}.

  \paragraph{Step 4: dynamical survival of spectral sectors.}
    The subsequent step is to simulate the full $Q_8$ substrate with the DBI
    constraint imposed locally at each vertex.
    The modal decomposition
    \[
      A_n(t) = a_n(t)\|\psi_n\|_\infty
    \]
    can then be tracked dynamically.
    The key question is whether spectral cascades concentrate energy in the
    sector with the largest admissible window, namely the non-abelian
    $\rho_4$ sector.

  \paragraph{Step 5: extension to larger graphs.}
    Finally, the toy model should be extended to Cayley graphs of larger
    groups approximating $S^3$ more faithfully.
    Natural candidates include the binary icosahedral group $2I$ (order~120)
    and families of Cayley graphs on $\mathrm{SL}(2,\mathbb{F}_p)$.

    In particular, Ramanujan graph constructions such as the
    Lubotzky--Phillips--Sarnak graphs provide sequences with optimal spectral
    gap properties.
    Studying bounded-flux filtering on such graphs would test whether the
    spectral admissibility hierarchy persists in large expanders and may
    reveal a deeper connection between substrate capacity constraints and
    the spectral theory of expander graphs.

    \bibliographystyle{ieeetr}
    \bibliography{refs}

\end{document}
