\subsection{Relation to existing approaches.}
  \label{subsec:relation-to-existing-approaches}
  The present construction shares motivations with several background-independent
  approaches to quantum gravity, including the causal set theory, loop quantum gravity,
  and spectral geometry.
  Unlike causal set models, no fundamental discreteness is postulated, and continuity is
  maintained at all levels, with effective granularity arising from spectral filtering
  rather than from an underlying lattice.
  In contrast to loop-based approaches, no spin networks or combinatorial structures
  have been introduced as fundamental entities.
  Compared to non-commutative or spectral geometry programs, the emphasis here is not
  on algebraic generalization of manifolds, but on operational reconstruction of metric
  structures from admissible relational correlations.

  \paragraph{Relation to graph Laplacian convergence results.}
    The spectral reconstruction developed in this work operates within a well-established
    analytical framework.
    The convergence of graph Laplacians toward the Laplace--Beltrami operator on a
    smooth manifold in the dense sampling limit was established by Belkin and
    Niyogi~\cite{Belkin2003} and refined by Hein, Audibert, and von
    Luxburg~\cite{Hein2007}.
    These results ensure that the spectral data of the relational graph faithfully
    approximate the continuous spectral data of the underlying geometric object as the
    graph becomes dense, under mild regularity assumptions on the sampling measure.

    The use of spectral proximity to define effective distances is closely related to the
    diffusion map framework of Coifman and Lafon~\cite{Coifman2006}, in which a
    diffusion distance is constructed from the heat kernel of the graph Laplacian.
    The spectral distance $d_S(i,j)$ introduced in Section~3 is closely related to the
    diffusion distance at fixed diffusion time, differing only by the choice of spectral
    weighting.
    The present framework differs from diffusion maps in two respects: the distance is
    used as an input to a hierarchical construction of the effective field
    $\chi_\mathrm{eff}$ (Appendix~B), rather than as a standalone embedding tool, and
    the admissibility filter $F_{\lambda^*}$ imposes an additional spectral regularity
    condition that has no direct counterpart in standard manifold learning.

    The analytical underpinning of geodesic reconstruction from spectral data traces
    back to Varadhan's formula~\cite{Varadhan1967}:
    \begin{equation}
      d(x, y)^2 = -4t \log p_t(x, y) + o(t) \quad \text{as } t \to 0^+,
    \end{equation}
    where $p_t(x,y)$ is the heat kernel of the manifold.
    This asymptotic relation establishes that short-time spectral data encode the
    Riemannian geodesic distance, providing the analytical foundation for spectral
    reconstruction methods.
    The spectral distances used here exploit the same correspondence in the finite
    resolution regime, replacing the continuous heat kernel by its discrete spectral
    approximation.

    The approach differs structurally from Connes' noncommutative
    geometry~\cite{Connes1994}, which reconstructs a metric from a spectral triple
    $(\mathcal{A}, \mathcal{H}, D)$ involving a Dirac operator.
    The present framework uses only the scalar Laplacian and does not require a spinor
    structure, an algebra of observables, or a specific operator-theoretic axiom set.
    The present construction may be interpreted as a restricted scalar instance of
    spectral geometry, but it does not rely on the full operator-algebraic framework of
    noncommutative geometry.
    The two approaches share the goal of recovering geometric structure from spectral
    data but operate at different levels of generality and with different motivating
    physical assumptions.
