\subsection{Scope and limitations.}
  \label{subsec:scope-and-limitations.}
  The present study is intentionally restricted to the emergence of geometric and gravitational structures.
  No attempt has been made to account for the matter degrees of freedom, energetic
  quantities, quantum statistics, or particle properties.
  These notions presuppose additional layers of effective description that lie beyond
  the geometric regime addressed here.

  Similarly, strongly non-factorizable or highly constrained configurations, for which
  no stable geometric parametrization exists, fall outside the scope of the present analysis.
  In such regimes, spacetime ceases to be an appropriate descriptive language,
  and alternative relational characterizations must be employed.

  \paragraph{Kinematic status and constraints on admissible dynamics.}
    The framework developed in this work is deliberately kinematic: no equations of
    motion, Lagrangian, or Hamiltonian are postulated for the relational substrate $\chi$.
    This restriction is intentional.
    The aim is to identify the structural conditions under which an effective geometric
    description is admissible, independently of the specific dynamical law governing
    the evolution of $\chi$.

    This kinematic character does not render the framework non-predictive.
    On the contrary, the spectral invariants identified here --- in particular the ratio
    $\lambda_2/\lambda_1 = 8/3$ in four-dimensional projectable regimes --- impose
    necessary constraints on any compatible dynamics.
    Any dynamical law governing $\chi$ that preserves the admissibility conditions of
    Section~4.1 must reproduce these spectral invariants in the geometric regime, at least
    in the regime where a geometric interpretation is admissible.
    Conversely, a dynamical model whose solutions generically violate the spectral
    admissibility conditions cannot be considered a consistent completion of the present
    kinematic framework.

    The distinction between kinematic structure and dynamical completion is standard in
    foundational physics.
    General relativity first specifies the geometric class of Lorentzian manifolds,
    while the Einstein equations act as a dynamical selection principle within that class.
    The present work occupies an analogous position at the level of the relational
    substrate: it identifies the admissible geometric regime without prescribing the
    dynamics that selects or sustains it.
