\subsection{Analytical derivation of the spectral ratio on $S^3$}
  \label{sec:analytical_ratio}

  The numerical convergence of the ratio $\lambda_2 / \lambda_1$ toward $8/3$ observed in
  the Monte--Carlo simulations of Appendix~D admits an exact analytical counterpart in the
  continuous setting, which we now establish.

  \paragraph{Spectrum of the scalar Laplacian on $S^3$.}
    Here $S^3$ is taken with unit radius and $\Delta_{S^3}$ denotes the positive
    Laplace--Beltrami operator $\Delta_{S^3} = -\operatorname{div}\operatorname{grad}$.
    The eigenvalues of $\Delta_{S^3}$ are given by the standard formula
    (see, e.g., \cite{Berger1971})
    \begin{equation}
      \label{eq:S3spectrum}
      \lambda_k = k(k + 2), \qquad k = 0, 1, 2, \ldots,
    \end{equation}
    with multiplicity $(k+1)^2$.
    The first non-trivial eigenvalue ($k=1$) is therefore $\lambda_1 = 3$,
    and the second distinct eigenvalue ($k=2$) is $\lambda_2 = 8$.
    Their ratio is exactly
    \begin{equation}
      \label{eq:exact_ratio}
      \frac{\lambda_2}{\lambda_1} = \frac{8}{3}.
    \end{equation}
    This result follows directly from the representation theory of
    $\mathrm{SU}(2) \cong S^3$: the eigenspaces are irreducible representations of
    $\mathrm{SU}(2)$ labeled by $k$, so the spectrum is algebraically determined by the
    group structure and is independent of any choice of coordinates or discretization.

  \paragraph{Role of the Hopf fibration.}
    The Hopf fibration $S^1 \hookrightarrow S^3 \to S^2$ provides a geometric decomposition
    of $S^3$ into base and fiber directions, but it does not alter the intrinsic spectrum of
    $\Delta_{S^3}$.
    The first non-trivial eigenvalue ($k=1$) corresponds to the fundamental irreducible
    representation of $\mathrm{SU}(2)$, and the ratio $\lambda_2/\lambda_1 = 8/3$ follows
    purely from this representation structure.
    The fibration offers a natural interpretation of the $k=1$ modes in terms of base and
    fiber excitations, but the ratio itself is an intrinsic spectral property of the
    three-sphere, not an artifact of the fibered construction.

    In the present framework, the effective four-dimensional character arises once the $S^1$
    fiber is interpreted as a compact gauge direction, yielding a three-dimensional base
    geometry supplemented by an internal degree of freedom.
    This interpretation is consistent with the fibered graph construction used in the
    numerical simulations of Appendix~D, where the $k$-NN graph is sampled from $S^3$ with
    the Hopf structure made explicit through the anisotropic kernel $K_\alpha$.

  \paragraph{Relation to the numerical estimator.}
    The Monte--Carlo and spectral estimators in Appendix~D compute the ratio
    $R(\alpha) = E_\mathrm{fiber}(\alpha)/E_\mathrm{base}(\alpha)$ of kernel-weighted edge
    energies along fiber and base directions respectively.
    In the isotropic regime ($\alpha \leq 0$), this ratio stabilizes at $R_0 \simeq 0.876$,
    reflecting the geometric partition of $S^3$ into fiber and base contributions under
    isotropic sampling.
    As $\alpha$ increases and fiber modes are selectively excited, the normalized quantity
    $E_\mathrm{fiber}(\alpha)/E_\mathrm{fiber}(0)$ converges to $8/3$.
    This convergence is consistent with the exact result~\eqref{eq:exact_ratio}: in the
    limit of strong fiber excitation, the ratio of fiber to base spectral energy approaches
    the ratio of the corresponding eigenvalues, which equals $8/3$ by the spectral
    formula~\eqref{eq:S3spectrum}.
    The two quantities --- $R_0$ and $8/3$ --- measure different aspects of the same
    spectral structure and are therefore not in contradiction.

  \paragraph{Diagnostic role.}
    The ratio $\lambda_2/\lambda_1 = 8/3$ is a \emph{topological spectral invariant}
    of $S^3$, not a universal constant of all relational graphs.
    Its role in the present framework is diagnostic: a relational substrate exhibiting this
    ratio in its projectable regime is spectrally compatible with a geometry modeled on $S^3$,
    and the Hopf structure provides a natural interpretation in terms of a four-dimensional
    spacetime with a compact gauge fiber.
    Substrates with different topologies yield different ratios and are thereby
    distinguishable at the spectral level without reference to any background metric.
    This discriminating property makes the ratio a stronger structural signature than
    metric reconstruction alone, since many distinct relational configurations may project
    to indistinguishable effective metrics while still being separated by their spectral
    invariants.
