\section{Relation to the Effective Geometric Description}
  \label{sec:relation-to-the-effective-geometric-description}

  The effective geometric structures introduced in the main text --- such as metric fields,
  spatial gradients, connection-like objects, and Poisson-type equations --- do not
  represent fundamental degrees of freedom in the present relational framework.
  They arise as effective descriptions of observable relational data under
  projection, once a projectable regime becomes applicable.

  In the pre-geometric formulation, the relational substrate is defined purely in terms
  of adjacency, spectral properties, and admissibility constraints, without reference to
  coordinates, distances, or differential structures.
  Geometric notions become meaningful only after a stable spectral continuum limit
  exists and an operational distance $d^W$ has been constructed from relational stiffness
  (Appendix~\ref{app:relational-distance-minimal-path}).

  Within this regime, smooth variations of the effective spectral description over
  neighbourhoods defined by $d^W$ admit a continuum approximation.
  Metric components, gradients, and connection-like quantities were then introduced
  as \textit{descriptive tools} that compactly encode how admissible relational
  correlations respond to local perturbations.
  They summarize the collective response properties of the projected description
  rather than encoding independent dynamic degrees of freedom.

  Importantly, these geometric objects are valid only insofar as the projection remains
  locally injective and relational variations remain weak.
  When admissibility breaks down near the deprojection thresholds or in strongly
  constrained regions, the effective geometric description loses its operational meaning.
  In such regimes, no failure of geometric dynamics is implied; rather, the geometric
  language itself ceases to be applied.

  Therefore, the effective geometric description employed in the main text should
  be understood as a regime-dependent and operational representation of relational
  organization, exactly within its domain of validity and silent outside it.
