\subsection{The Geometric Effective Description of $\chi$ Dynamics}
  \label{subsec:geometric-action}

  \subsubsection*{Interpretational caution.}
    The following discussion introduces effective geometric language (such as ``metric'',
    ``distance'', or ``spacetime'') solely as a descriptive tool to characterize the relational
    structure of $\chi$ configurations in regimes where a quasi-stable geometric interpretation
    emerges.
    These are not fundamental ingredients of the theory, but derived constructs valid only when
    $\chi$ admits a smooth, macroscopic description.

  \subsubsection{Effective Observables from $\chi$ Correlations}
    In Cosmochrony, geometric notions---time intervals, spatial separations, and causal
    structure---emerge from the relational properties of $\chi$ configurations.
    No background spacetime or discrete structure is assumed.
    Instead, observables are constructed directly from $\chi$ and its local variation structure.

    Operational time intervals are defined by the accumulated relaxation of $\chi$ along a path
    in configuration space:
    \begin{equation}
      \tau_{AB} \propto \int_{\gamma_{AB}} \mathcal{D}_{\lambda} \chi \, d\lambda,
    \end{equation}
    where $\gamma_{AB}$ is a path connecting configurations $A$ and $B$, and
    $\mathcal{D}_{\lambda}\chi$ is the relaxation rate of $\chi$.

    Operational spatial separation is quantified by the decay of $\chi$ correlations between two
    configurations:
    \begin{equation}
      d(x,y) \propto -\log\left(\frac{\langle \chi(x)\chi(y)\rangle}{\langle \chi^2\rangle}\right),
    \end{equation}
    where $\langle \chi(x)\chi(y)\rangle$ measures the correlation between configurations labeled
    by $x$ and $y$.
    These definitions are purely relational and make no reference to a pre-existing metric or
    discrete structure.
    They are applicable whenever $\chi$ configurations exhibit stable patterns.

  \subsubsection{Effective Metric as a Descriptive Tool}
    In regimes where $\chi$ varies smoothly, the observables above can be summarized by an
    effective metric tensor $g_{\mu\nu}[\chi]$.
    This metric is not a fundamental object, but a compact representation of how $\chi$
    correlations modulate causal connectivity and proper intervals.

    The metric is introduced as a descriptive summary of $\chi$ correlations, not postulated as
    an independent degree of freedom.
    For example:
    \begin{itemize}
      \item The conformal (lightcone) structure is constrained by the maximal relaxation speed
      $c$ of $\chi$.
      \item Proper time between effective events is proportional to the accumulated $\chi$
      relaxation along paths connecting the corresponding configurations.
      \item Spatial distance reflects the decay rate of $\chi$ correlations.
    \end{itemize}

    No discrete-to-continuum limit is invoked.
    The theory is continuous at all scales; apparent granularity (e.g., Planck-scale phenomena)
    is attributed to non-linear $\chi$ dynamics rather than to an underlying discretization.

    No background $\eta_{\mu\nu}$ is assumed.
    Minkowski space appears only as a convenient approximation in suitable limits (e.g.,
    weak-gradient regimes), without ontological status in the fundamental description.

  \subsubsection{Consistency with General Relativity}
    The effective metric $g_{\mu\nu}[\chi]$ constructed as a summary of $\chi$ correlations
    reproduces the phenomenology of general relativity in the following sense:
    \begin{itemize}
      \item \textbf{Weak-field limit:} when $\chi$ gradients are small, the effective metric
      approaches a form compatible with Einstein-like dynamics for a fluid-like
      stress-energy description associated with $\chi$ excitations.
      \item \textbf{Strong-field regimes:} near localized $\chi$ excitations (e.g., solitons), the
      metric encodes time dilation and spatial curvature as emergent effects of slowed $\chi$
      relaxation, without requiring a fundamental gravitational field.
      \item \textbf{Cosmological expansion:} homogeneous relaxation of $\chi$ yields an effective
      Hubble-like expansion law for the emergent scale factor.
    \end{itemize}

    Crucially, this is not a bootstrap procedure.
    The metric is not iteratively reconstructed from $\chi$; it is a post-hoc descriptive tool
    summarizing geometric regularities of $\chi$ configurations.
    The theory's predictive content resides in the dynamics of $\chi$, not in the metric itself.

  \subsubsection{Ontological Status of the Metric}
    To avoid confusion, we emphasize:
    \begin{itemize}
      \item $\chi$ is the only fundamental field.
      Spacetime, metric structure, and matter are emergent descriptions of $\chi$
      configurations.
      \item No ``double ontology'' is assumed: there is no underlying discrete graph or lattice.
      Geometric language is introduced only as an effective tool.
      \item The metric $g_{\mu\nu}[\chi]$ is an effective construct, analogous to how temperature
      emerges in thermodynamics.
      It is useful for coarse-grained description but plays no role in the fundamental
      dynamics.
    \end{itemize}

  \subsubsection{Summary: A Fully Continuous Framework}
    \begin{itemize}
      \item \textbf{Fundamental level:} only the continuous $\chi$ field exists, evolving through
      a monotonic relaxation dynamics.
      \item \textbf{Effective level:} geometric observables (time, distance, metric) emerge from
      $\chi$ correlations in suitable regimes.
      \item \textbf{No bootstrap:} the metric is never iteratively constructed or assumed; it is a
      derived description summarizing the relational structure of $\chi$ configurations.
      \item \textbf{No discretization:} apparent ``Planck-scale'' effects arise from non-linear
      $\chi$ dynamics, not from an underlying discrete substrate.
    \end{itemize}
