\section{Discussion}
  \label{sec:discussion}

  The present work has developed a relational and spectral route to emergent metric
  geometry, deliberately avoiding the introduction of spacetime structure as a
  fundamental postulate.
  Instead, geometric notions arise as effective descriptors of admissible relational
  configurations selected by spectral filtering and projection.

  A central result is that metricity can be reconstructed operationally from purely
  relational spectral data, without assuming a background manifold, predefined
  distance function, or fundamental locality.
  The effective metric appears as a compact summary of correlation structure in regimes
  where admissible configurations admit a stable and approximately factorizable
  description.

  \input{06-discussion/01-structural-robustness}

  \paragraph{Status of spacetime geometry.}
    Within this framework, spacetime geometry is neither fundamental nor dynamical in the
    usual sense.
    It functions as an emergent kinematical structure, applicable only when the
    projection from the relational substrate becomes locally injective and spectrally
    well-conditioned.
    Outside these regimes, geometric notions lose operational meaning rather than being
    modified or replaced by alternative field equations.

    This perspective clarifies the remarkable universality of general relativity.
    Whenever a smooth geometric description exists and admissible configurations vary
    slowly compared to the spectral cutoff, the standard geometric relations of general
    relativity are necessarily recovered.
    Einstein’s equations thus appear not as a microscopic law of the substrate, but as a
    consistency condition governing emergent geometry wherever spacetime itself is a
    valid descriptor.

    A more detailed clarification of the representational status of the effective
    geometric description, and its relation to the underlying relational formulation,
    is provided in Appendix~\ref{subsec:relation-to-the-effective-geometric-description}.

  \paragraph{Relation to existing approaches.}
    The present construction shares motivations with several background-independent
    approaches to quantum gravity, including causal set theory, loop quantum gravity,
    and spectral geometry.
    Unlike causal set models, no fundamental discreteness is postulated; continuity is
    maintained at all levels, with effective granularity arising from spectral filtering
    rather than from an underlying lattice.
    In contrast to loop-based approaches, no spin networks or combinatorial structures
    are introduced as fundamental entities.
    Compared to non-commutative or spectral geometry programs, the emphasis here is not
    on algebraic generalization of manifolds, but on operational reconstruction of metric
    structure from admissible relational correlations.

  \paragraph{Non-uniqueness and regime dependence.}
    An important implication of the spectral approach is the intrinsic non-uniqueness of
    geometric reconstruction.
    Different relational configurations may give rise to indistinguishable effective
    metrics within a given regime, reflecting the non-injective character of the
    projection.
    Geometric descriptions are therefore intrinsically approximate and regime-dependent.
    This non-uniqueness should not be interpreted as an ambiguity of the theory, but as a
    structural feature of emergent geometry itself.

  \paragraph{Spectral rigidity and structural invariants.}
    While the reconstruction of effective geometry from relational spectral data is
    intrinsically non-unique and regime-dependent, this non-uniqueness does not extend
    to all spectral observables.
    The relational substrate exhibits robust spectral invariants that persist across
    discretizations, graph realizations, and numerical schemes.

    A notable example is the emergence of the universal ratio
    $\lambda_2/\lambda_1 = 8/3$ in the scalar Laplacian spectrum, as illustrated in
    Appendix~\ref{subsec:emergent-coordinates}.
    Although many relational microstates may project to indistinguishable effective
    metrics, only a restricted class of spectral organizations is compatible with a
    stable projectable geometric regime.

    Such invariants therefore constrain the space of admissible emergent geometries
    more strongly than metric reconstruction alone.
    They provide an intermediate level of structure between microscopic relational
    descriptions and macroscopic geometric observables, and may serve as distinguishing
    signatures with respect to other background-independent approaches that reproduce
    similar effective geometries without exhibiting comparable spectral constraints.

    In this sense, the existence of robust spectral invariants such as the
    $\lambda_2/\lambda_1 = 8/3$ ratio suggests that relational substrates may be
    more tightly constrained than effective metric reconstructions alone would
    indicate.

  \paragraph{Scope and limitations.}
    The present work is intentionally restricted to the emergence of geometric and
    gravitational structure.
    No attempt has been made to account for matter degrees of freedom, energetic
    quantities, quantum statistics, or particle properties.
    These notions presuppose additional layers of effective description that lie beyond
    the geometric regime addressed here.

    Similarly, strongly non-factorizable or highly constrained configurations, for which
    no stable geometric parametrization exists, fall outside the scope of the present
    analysis.
    In such regimes, spacetime itself ceases to be an appropriate descriptive language,
    and alternative relational characterizations must be employed.

  \paragraph{Outlook.}
    By isolating the emergence of metric geometry from other physical structures, this
    work establishes a clean and minimal foundation for subsequent investigations.
    Extensions of the framework to non-geometric regimes, to the emergence of matter-like
    excitations, or to quantum and statistical phenomena require additional assumptions
    and will be addressed separately.

    The results presented here demonstrate that a large class of gravitational phenomena
    can be understood as consequences of relational spectral organization alone.
    In this sense, geometry appears not as the stage on which physics unfolds, but as a
    derived and context-dependent summary of deeper relational structure.
