\subsection{Static Spherically Symmetric Geometry}
  \label{subsec:schwarzschild-recovery}

  We now consider a static and approximately spherically symmetric relational
  configuration that allows a smooth continuum description.
  No dynamical assumptions or field equations were introduced.
  The analysis is purely kinematic and focuses on a geometric form compatible
  with symmetry and weak-field consistency.

  In such a regime, the effective metric can be written in a standard static
  spherically symmetric form,
  \begin{equation}
    ds^2 = -A(r)\,c^2 dt^2 + B(r)\,dr^2 + r^2 d\Omega^2 ,
  \end{equation}
  where the metric functions $A(r)$ and $B(r)$ encode the operational relationship
  between the temporal and spatial intervals.

  In the weak-field limit, the metric coefficients admit an expansion
  \begin{equation}
    A(r) \simeq 1 + 2\frac{\Phi(r)}{c^2},
    \qquad
    B(r) \simeq \left(1 + 2\frac{\Phi(r)}{c^2}\right)^{-1},
  \end{equation}
  where $\Phi(r)$ is the effective potential for characterizing the deviation from a flat geometry.
  This definition is purely geometric and does not presuppose any dynamic origin
  for $\Phi$.

  Requiring asymptotic flatness and consistency with spherical symmetry uniquely
  fixes the exterior form of the metric to the leading order,
  \begin{equation}
    A(r) = 1 - \frac{r_s}{r},
    \qquad
    B(r) = \left(1 - \frac{r_s}{r}\right)^{-1},
  \end{equation}
  where $r_s$ is an integration constant with length dimensions.
  This geometry coincides with the Schwarzschild metric in leading order.

  The emergence of this form does not rely on a postulated gravitational field equation.
  This reflects the fact that, among static spherically symmetric geometries, the
  Schwarzschild metric provides a unique weak-field extension compatible with
  asymptotic flatness and local isotropy.

  Standard weak-field tests follow directly from the geometric structures.
  Gravitational redshift, light deflection, and time dilation arise as kinematic
  consequences of the metric coefficients and do not require any additional
  assumptions.
