\subsection{Relational Structure}
  \label{subsec:relational-structure}

  We introduce a minimal relational framework intended to capture the weakest
  structural assumptions required for spectral reconstruction of geometry.
  No spacetime manifold, metric tensor, coordinate system, or causal structure is
  postulated at this level.
  The framework is deliberately formulated in non-geometric terms.

  The system is described by a set of abstract relational elements together with
  admissible relations among them.
  These relations define a connectivity structure, which may be represented
  discretely or in a coarse-grained form.
  No embedding space or background geometry is assumed or required.

  The relational elements themselves are not assigned local values, dimensional
  quantities, or order parameters.
  Such quantities arise only at an effective level, when relational configurations
  admit a stable spectral reconstruction.
  In particular, notions of distance, dimension, and curvature are not fundamental
  primitives but reconstructed descriptors.

  A key feature of the reconstruction problem is its inherent non-injectivity:
  distinct relational configurations may give rise to identical effective geometric
  descriptions, while a single configuration may admit multiple equivalent spectral
  representations.
  This loss of information is a structural property of spectral reconstruction and
  does not rely on any physical coarse-graining mechanism.

  Throughout this work, geometric and field-theoretic language is employed strictly
  as an effective descriptive convenience.
  Such language refers to reconstructed structures that become meaningful only in
  regimes where the relational system supports a consistent spectral continuum
  approximation.
  It does not imply the existence of underlying spacetime entities or dynamical
  fields.

  This relational starting point provides a neutral basis for defining spectral
  operators and for investigating how effective geometric properties can emerge
  from connectivity data alone.
