\section{Continuum Limit and Emergent Metric}
  \label{sec:continuum-limit-and-emergent-metric}

  The spectral distance introduced in the previous section provides a purely relational notion of the proximity.
  A natural question is under what conditions such a construction admits an effective
  continuum description, allowing the use of familiar geometric tools.
  In this section, we analyze the regimes in which a relational spectral structure can be
  approximated using a smooth metric geometry.

  We consider families of relational graphs or coarse-grained networks whose Laplacian
  spectra exhibit sufficient regularity.
  No assumptions were made concerning the underlying manifold or dimensionality.
  Instead, the continuum limit is operationally defined, as the regime in which spectral
  distances vary smoothly and admit consistent local approximations.

  A key requirement for the emergence of continuum description is local factorizability.
  In such regimes, the relational system admits an approximate decomposition into
  weakly coupled subsystems, allowing the spectral distances to be defined locally and independently.
  This property underlies the emergence of an effective locality and permits the
  introduction of local coordinate charts as auxiliary descriptive tools.

  The continuum approximation is inherently non-unique.
  Because spectral reconstruction is generically non-injective, distinct relational
  configurations may give rise to identical effective metric descriptions.
  Conversely, a given relational structure can admit multiple equivalent continuum embeddings.
  This ambiguity is a structural feature of the spectral geometry and does not signal a loss
  of consistency.

  When local factorizability and spectral regularity are satisfied, spectral distance
  admits a local quadratic approximation.
  In this regime, one may introduce an effective metric tensor $g_{\mu\nu}(x)$ is defined by
  the leading-order expansion:
  \begin{equation}
    d_{\mathrm{spec}}(i,j)^2
    \;\approx\;
    g_{\mu\nu}(x)\,\Delta x^\mu \Delta x^\nu ,
  \end{equation}
  where $\Delta x^\mu$ denotes the coordinate differences in the local embedding.
  The metric tensor is not a fundamental structure but a derived descriptor that encodes the
  local behavior of spectral distances.

  The validity of continuum approximation is limited to regimes in which higher-order
  spectral corrections remain subdominant.
  Outside such regimes, no smooth geometric interpretation is assumed, and the relational
  description must be treated in its full spectral form.

  The framework developed in this study is purely kinematic.
  No assumptions regarding dynamics, temporal ordering, or causal structure were required
  for the emergence of an effective metric description.
  The analysis focuses solely on the conditions under which the relational spectral data
  support a continuum geometric approximation.

  \input{04-continuum-limit-and-emergent-metric/01-spectral-admissibility}

  In some reconstruction schemes, admissibility criteria may involve monotonic
  spectral filters or ordering parameters, reflecting the fact that effective
  continuum descriptions are typically insensitive to fine-grained spectral
  rearrangements beyond a given resolution.

  We emphasize that the spectral reconstruction of an effective metric does not
  require the introduction of additional fundamental fields or degrees of freedom;
  the geometric structures discussed here arise solely from relational spectral
  data and their continuum approximation.
