\subsection{Validity of Geometric Descriptions}
  \label{subsec:validity-of-geometry}

  The emergence of curvature and metric structure discussed in the previous sections
  does not imply that geometric descriptions are universally applicable.
  Rather, geometry appears to be an effective language, valid only in regimes where the
  relational structure admits a smooth and locally stable continuum approximation.

  In such regimes, geometric relations exhibit a remarkable robustness.
  The local curvature, geodesic deviation, and horizon structure depend only weakly on
  the microscopic details of the underlying relational system.
  This explains the universality of geometric behavior observed across a wide range of physical situations.

  Outside these regimes, the failure of the geometric descriptions should not be
  interpreted as a breakdown of the geometric laws.
  Instead, it reflects the loss of applicability of the geometric language.
  When local injectivity or smoothness conditions are violated, the notion of
  spacetime ceases to be operationally meaningful, and no geometric description is assumed.

  From this perspective, geometric relations function as internal consistency
  conditions for emergent spacetime descriptions rather than as fundamental laws governing an underlying substrate.
  Geometry is exactly where it is applied and silent elsewhere.

  \paragraph{Relational encoding of mass.}
    In the present framework, the mass parameter appearing in the Schwarzschild-type
    effective geometry is not introduced as a fundamental source term.
    Instead, it reflects a persistent and localized modification of the relational connectivity of the $\chi$ substrate.

    Operationally, a massive configuration corresponds to a region in which admissible
    relaxation paths are systematically constrained, resulting in an effective reduction in
    the relational accessibility across that domain.
    This constraint modifies the local spectral response of the relational Laplacian
    while preserving global admissibility conditions.
    In the projectable regime, this localized alteration is encoded geometrically as a
    radial deformation of the effective distances, whose asymptotic form reproduces the Schwarzschild mass parameter.

    The mass therefore quantifies the integrated obstruction to relational relaxation,
    rather than the presence of an external source.
    This interpretation is consistent with the non-fundamental status of the effective
    metric and clarifies how gravitational mass emerges from purely relational modifications of the substrate.
