\subsection{Entanglement and Nonlocal Correlations}
  \label{subsec:entanglement-nonlocality}

  Within the Cosmochrony framework, quantum entanglement does not correspond to a
  physical interaction or linkage between spatially separated entities.
  Instead, it reflects the persistence of a shared relational structure within a single, non-factorizable
  configuration of the $\chi$ substrate.

  Entangled systems arise when a unified relational configuration admits an effective
  projection onto spatially separated degrees of freedom.
  Although the projected descriptions assign distinct spacetime locations to the corresponding subsystems,
  these locations represent different local projections of one and the same underlying relational configuration.

  As a consequence, admissible projected descriptions cannot, in general, be decomposed into independent subsystems
  without loss of relational consistency.
  Measurement operations act on a globally defined relational structure whose admissible projections must remain
  mutually compatible.
  Once a particular effective description is selected in one region, the set of admissible descriptions elsewhere is
  correspondingly restricted, without any physical signal exchange or dynamical influence.

  In this sense, quantum nonlocality in Cosmochrony is ontological rather than dynamical.
  The underlying relational configuration is globally defined, while all effective evolution
  and reprojection processes remain governed locally by the bounded relaxation and
  projectability constraints of $\chi$.
  Entanglement correlations are therefore fully compatible with relativistic causality and do not require the
  introduction of nonlocal forces, preferred reference frames, or superluminal influences.

  A formal relational analysis of non-factorization and its role in entanglement is provided
  in Appendix~\ref{subsec:non-factorization-entanglement}.
