\subsection{Spectral Characterization of Mass and the Secondary Role of $V(\chi)$}
  \label{subsec:spectral_mass}

  This appendix clarifies the conceptual status of inertial mass in Cosmochrony.
  While mass originates physically from the resistance of solitonic configurations
  to the relaxation of the $\chi$ field, this resistance admits a quantitative
  characterization in terms of the spectral properties of an associated stability
  operator.

  In this context, spectral analysis does not redefine the physical origin of
  inertial mass—which remains the resistance of localized configurations to $\chi$
  relaxation—but provides a structured and potentially calculable description of
  this resistance.

  A central conjecture of the Cosmochrony framework is that particle masses are not
  fundamental parameters encoded in the nonlinear potential $V(\chi)$.
  Instead, they emerge as spectral properties of a relaxation operator defined on a
  relational substrate, which may be represented, for calculational purposes, by a
  discrete graph structure.

  \paragraph{Mass spectrum as eigenmodes of a relaxation operator.}
    Localized particle-like excitations are identified with normal modes of an
    effective Laplace--Beltrami operator acting on a graph $G(V,E)$,
    \begin{equation}
      \Delta_G \psi_n = -\lambda_n \psi_n ,
    \end{equation}
    where $\psi_n$ are eigenmodes characterizing the stability of localized
    configurations.
    The associated inertial masses are conjectured to scale as
    \begin{equation}
      m_n c^2 \propto \sqrt{\lambda_n},
    \end{equation}
    in agreement with the effective spectral relations introduced in
    Section~\ref{subsec:perspectives_mass_spectrum}.
    Physically, this scaling reflects the fact that inertial mass measures the
    characteristic frequency associated with the resistance of a localized
    configuration to the relaxation of the $\chi$ field.

  This relation is analogous to the emergence of discrete vibrational frequencies
    in bounded elastic systems, where spectral values are fixed by geometry and
    connectivity rather than by adjustable parameters.
    Within Cosmochrony, mass hierarchies are therefore interpreted as geometric and
    topological properties of the underlying relational structure.

    A decisive test of this conjecture would consist in computing the low-lying
    spectrum of $\Delta_G$ on large but finite networks with physically motivated
    connectivity rules.
    Even approximate agreement with observed mass ratios would strongly support the
    spectral origin of inertial mass and the non-fundamental role of $V(\chi)$.

  \paragraph{Spectral structure and separation of descriptive levels.}

    To avoid circular dependencies between geometry, dynamics, and emergent particle
    properties, Cosmochrony distinguishes three conceptual levels.
    This separation is essential to avoid any circular definition in which emergent
    geometric notions would feed back into the operator responsible for mass
    generation.

    At the fundamental level, inertial masses are associated with the spectral
    properties of a background-independent relaxation operator
    $\Delta_G^{(0)}$, defined by the intrinsic relational connectivity of the
    substrate.
    This operator is not tied to any specific spacetime geometry or instantaneous
    $\chi$ configuration and provides a stable spectral structure.

    At the emergent geometric level, coarse-grained configurations of $\chi$ give rise
    to effective notions of spacetime, including gravitational time dilation and
    cosmological expansion.
    These geometric effects influence propagation and interaction, but do not modify
    the underlying spectral operator responsible for mass generation.

    Finally, fast dynamical processes such as radiation, scattering, and decoherence
    correspond to interaction-induced redistributions of relaxation potential within
    the $\chi$ field.
    These processes affect observables without redefining the fundamental spectral
    structure.

  \paragraph{Residual role of the potential $V(\chi)$.}

    Within this spectral picture, the nonlinear potential $V(\chi)$ plays a secondary
    and effective role.
    It provides a local, coarse-grained description of localization mechanisms
    associated with low-lying spectral modes, but does not independently set the
    overall mass scale.
  Its admissible form is constrained by the requirement that it support stable
    solitonic configurations compatible with the spectral structure.

  \paragraph{Supporting perspectives.}
    Additional constraints arising from discrete symmetries, information-theoretic
    considerations, or dynamical consistency may further restrict admissible
    connectivity structures.
    However, these considerations remain secondary to the central spectral hypothesis
    and are not required for its internal coherence.

    Taken together, these arguments suggest that a substantial part of the
    explanatory burden for mass generation in Cosmochrony lies in the spectral
    properties of the underlying relaxation dynamics, with $V(\chi)$ serving as a
    derived effective descriptor rather than a fundamental source of mass.

    Extending this spectral characterization toward concrete mass predictions
    requires specifying the relaxation operator and its boundary conditions,
    particularly for composite solitonic sectors, as discussed in
    Section~\ref{subsec:perspectives_mass_spectrum}.
